### 2024-09-25 proposal/project_proposal.tex
started the data section

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% for space filling
\usepackage{lipsum}
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data

\title{Proposal: An Analysis of Hockey's Most Controversial Statistic}
\author{Melanie Desroches\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\paragraph{Introduction}
In hockey, a very common statistic that is reported is the plus/minus statistic. Plus/minus is calculated
by taking all the even strength goals that were scored by their team while they were on the ice and subtracting
it by the total number of goals that were scored by the other team while they were on the ice. The idea behind 
the number is that it will show if a player has an overall positive or negative contribution when they are playing. 
However, the plus/minus has a lot of drawbacks. For example, in ice hockey, players tend to play on "lines", meaning 
the same three forwards and the same two defencemen tend to play at the same time. Because of this, plus/minus is heavily 
influenced by who they play with. Some would argue that the plus/minus is more of a team statistic rather than an 
individual statistics. Many of the criticisms of this statistic comes from the fact that there are a lot of confounding 
variables that influence how the plus/minus is calculated. Because of these criticisms, alternatives to the plus/minus
statistic have been developed, such as COSRI or Fenwick.

The goal of this paper is to evaluate the effectiveness of the plus/minus statistic. Some of the information
I am trying to determine is if the plus/minus statistic can be used to determine how good a player is both on
offense and defense. I also want to see if the plus/minus has any validity being used as an individual
statistic or as a team statistic. Based on my observations, I will compare the plus/minus statistic to other
statistics that have a similar purpose such as CORSI or Fenwick.

\paragraph{Specific Aims}
Formulate your research question;
translate your research question into statistical/data science questions

Research questions:
Is the plus/minus statistic an effective way of evaluating a players offensive or defensive
contribution?
Is the plus/minus statistic an individual or team statistic?
Is the plus/minus statistic a better evaluation of a player than other contribution statistics
such as CORSI or Fenwick?

\paragraph{Data}
Hopefully, you have identified the data needed for your project. Give a
description about it.

I am using data from the National Hockey League (NHL). The data is collected from a website called
Hockey Reference. I am particularly interested in the advanced hockey statistics that are collected
for each player in the NHL.

\paragraph{Research Design and Methods}
What design or methods will you use?

Since the plus/minus statistic tends to deal with collinearity, a ridge regression might be more
appropriate for analysis.

\paragraph{Discussion}
What are the most challenge parts of the task?
What are the limitations of your work? What is your fall-back plan if
something unexpected happens?

\bibliography{../manuscript/refs}
\bibliographystyle{chicago}

\end{document}

### 2024-09-26 proposal/project_proposal.tex
worked on intro, aims, and data

@@
 \paragraph{Introduction}
-In hockey, a very common statistic that is reported is the plus/minus statistic. Plus/minus is calculated
-by taking all the even strength goals that were scored by their team while they were on the ice and subtracting
-it by the total number of goals that were scored by the other team while they were on the ice. The idea behind 
-the number is that it will show if a player has an overall positive or negative contribution when they are playing. 
-However, the plus/minus has a lot of drawbacks. For example, in ice hockey, players tend to play on "lines", meaning 
-the same three forwards and the same two defencemen tend to play at the same time. Because of this, plus/minus is heavily 
-influenced by who they play with. Some would argue that the plus/minus is more of a team statistic rather than an 
-individual statistics. Many of the criticisms of this statistic comes from the fact that there are a lot of confounding 
-variables that influence how the plus/minus is calculated. Because of these criticisms, alternatives to the plus/minus
-statistic have been developed, such as COSRI or Fenwick.
-
-The goal of this paper is to evaluate the effectiveness of the plus/minus statistic. Some of the information
-I am trying to determine is if the plus/minus statistic can be used to determine how good a player is both on
-offense and defense. I also want to see if the plus/minus has any validity being used as an individual
-statistic or as a team statistic. Based on my observations, I will compare the plus/minus statistic to other
-statistics that have a similar purpose such as CORSI or Fenwick.
+In hockey, a very common yet controversial statistic that is reported is the plus/minus statistic. Plus/minus is 
+calculated by taking all the even strength goals that were scored by their team while they were on the ice and 
+subtracting it by the total number of goals that were scored by the other team while they were on the ice. The 
+idea behind the number is that it will show if a player has an overall positive or negative contribution when 
+they are playing. However, the plus/minus has a lot of drawbacks. For example, in ice hockey, players tend to play 
+on "lines", meaning the same three forwards and the same two defencemen tend to play at the same time. Because of this, 
+plus/minus is heavily influenced by who plays with who. Because of this, some argue that the plus/minus is more of a team 
+statistic rather than an individual statistic. Many of the criticisms of this statistic comes from the fact that
+there are a lot of confounding variables that influence how the plus/minus is calculated. Alternatives to the plus/minus 
+statistic have been developed, such as COSRI or Fenwick. Understanding the limitations and potential of plus/minus can 
+help coaches, analysts, and fans better interpret player data. While advanced statistics are valuable, plus/minus is still 
+popular, especially among casual fans. The project explores whether the plus/minus still holds relevance or whether it should 
+be supplemented or replaced with other metrics.
 
@@
 
-Research questions:
-Is the plus/minus statistic an effective way of evaluating a players offensive or defensive
-contribution?
-Is the plus/minus statistic an individual or team statistic?
-Is the plus/minus statistic a better evaluation of a player than other contribution statistics
-such as CORSI or Fenwick?
+The goal of this project is to evaluate the effectiveness of the plus/minus statistic. Some of the information
+I am trying to determine is if the plus/minus statistic can be used to determine how good a player is both on
+offense and defense. I also want to see if the plus/minus has any validity being used as an individual
+statistic or as a team statistic. Based on my observations, I will compare the plus/minus statistic to other
+statistics that have a similar purpose such as CORSI or Fenwick.
+These questions are critical as plus/minus is used in scouting and analysis, but its reliability 
+has been debated. By comparing it with modern metrics, I aim to shed light on its true value in player evaluation.
 
 \paragraph{Data}
-Hopefully, you have identified the data needed for your project. Give a
-description about it.
 
-I am using data from the National Hockey League (NHL). The data is collected from a website called
-Hockey Reference. I am particularly interested in the advanced hockey statistics that are collected
-for each player in the NHL.
+I am using data from the National Hockey League (NHL). The data is collected from a website called Natural Stat Trick. 
+I collected data from the 2021-2022, 2022-2023, and 2023-2024 seasons. I am particularly interested in the advanced hockey 
+statistics that are collected for each player in the NHL. These statistics include CORSI, Fenwick, high-danger chances,
+shooting percentages, and more.
 
@@
 Since the plus/minus statistic tends to deal with collinearity, a ridge regression might be more
-appropriate for analysis.
+appropriate for analysis. Similar work has been done in \cite{Macdonald_2012}
 

### 2024-09-26 proposal/project_proposal.tex
small changes to data and added to research

@@
 \paragraph{Specific Aims}
-Formulate your research question;
-translate your research question into statistical/data science questions
-
 The goal of this project is to evaluate the effectiveness of the plus/minus statistic. Some of the information
@@
 
-I am using data from the National Hockey League (NHL). The data is collected from a website called Natural Stat Trick. 
+I am using data from the National Hockey League (NHL). The data is collected from the websites Natural Stat Trick and NHLEdge. 
 I collected data from the 2021-2022, 2022-2023, and 2023-2024 seasons. I am particularly interested in the advanced hockey 
 statistics that are collected for each player in the NHL. These statistics include CORSI, Fenwick, high-danger chances,
-shooting percentages, and more.
+shooting percentages, and more. This data will also include more basic statistics such as points, goals, assists, plus/minus,
+time on ice, etc.
 
 \paragraph{Research Design and Methods}
-What design or methods will you use?
+The first step of the project is to collect all relevant data. In this case, I will be exporting the data from all players
+from the 2021- 2024 seasons into an Excel or csv file. Once the data has been collected, it needs to be cleaned. After the
+data has been prepared, I can start some initial analysis to see what variables are correlated and where issues of collineairty 
+exist, since this will effect which models and variables I use.
 

### 2024-09-27 proposal/project_proposal.tex
progress in research section and started discussion

@@
 statistic rather than an individual statistic. Many of the criticisms of this statistic comes from the fact that
-there are a lot of confounding variables that influence how the plus/minus is calculated. Alternatives to the plus/minus 
-statistic have been developed, such as COSRI or Fenwick. Understanding the limitations and potential of plus/minus can 
-help coaches, analysts, and fans better interpret player data. While advanced statistics are valuable, plus/minus is still 
-popular, especially among casual fans. The project explores whether the plus/minus still holds relevance or whether it should 
-be supplemented or replaced with other metrics.
+there are a lot of confounding variables that influence how the plus/minus is calculated. According to Brian MacDonald 
+"The traditional plus-minus statistic in hockey is highly dependent on a player’s teammates and opponents" \cite{Macdonald_2011}.
+Alternatives to the plus/minus statistic have been developed, such as COSRI or Fenwick. Understanding the limitations and
+potential of plus/minus can help coaches, analysts, and fans better interpret player data. While advanced statistics are valuable, 
+plus/minus is still popular, especially among casual fans. The project explores whether the plus/minus still holds relevance or 
+whether it should be supplemented or replaced with other metrics.
 
 \paragraph{Specific Aims}
-The goal of this project is to evaluate the effectiveness of the plus/minus statistic. Some of the information
-I am trying to determine is if the plus/minus statistic can be used to determine how good a player is both on
-offense and defense. I also want to see if the plus/minus has any validity being used as an individual
-statistic or as a team statistic. Based on my observations, I will compare the plus/minus statistic to other
-statistics that have a similar purpose such as CORSI or Fenwick.
-These questions are critical as plus/minus is used in scouting and analysis, but its reliability 
-has been debated. By comparing it with modern metrics, I aim to shed light on its true value in player evaluation.
+The goal of this project is to evaluate the effectiveness of the plus/minus statistic. Some of the information I am trying
+to determine is if the plus/minus statistic can be used to determine how good a player is both on offense and defense. I 
+also want to see if the plus/minus has any validity being used as an individual statistic or as a team statistic. Based on my 
+observations, I will compare the plus/minus statistic to other statistics that have a similar purpose such as CORSI or Fenwick. 
+These questions are critical as plus/minus is used in scouting and analysis, but its reliability has been debated. By comparing 
+it with modern metrics, I aim to shed light on its true value in player evaluation.
 
@@
 \paragraph{Research Design and Methods}
+This study will evaluate the effectiveness of the plus/minus statistic in ice hockey to determine whether it can assess a 
+player's offensive or defensive contributions and if it should be treated as an individual or team statistic. The analysis 
+will also compare plus/minus with other advanced hockey metrics, such as Corsi and Fenwick, to assess its overall utility.
+
 The first step of the project is to collect all relevant data. In this case, I will be exporting the data from all players
 from the 2021- 2024 seasons into an Excel or csv file. Once the data has been collected, it needs to be cleaned. After the
-data has been prepared, I can start some initial analysis to see what variables are correlated and where issues of collineairty 
+data has been prepared, I can start some initial analysis to see what variables are correlated and where issues of collinearity 
 exist, since this will effect which models and variables I use.
 
-Since the plus/minus statistic tends to deal with collinearity, a ridge regression might be more
-appropriate for analysis. Similar work has been done in \cite{Macdonald_2012}
+In order to determine if or how plus/minus can be used to assess offensive and defensive contribution, I will use a ridge 
+regression. Since the plus/minus statistic tends to deal with collinearity, a ridge regression might be more appropriate for 
+analysis. Similar work has been done in \cite{Macdonald_2012}. The coefficients from the ridge regression model will reveal 
+whether offensive or defensive metrics contribute more to explaining plus/minus. I plan on using a similar
 
@@
 
+According to Brian MacDonald in his ridge regression analysis of the plus/minus, "Ridge regression frequently reduces the error 
+bounds in the estimates and improves the predictive performance of the model when collinearly exists in the data. Ridge regression 
+introduces bias in the estimates, but the tradeoff is typically worthwhile." \cite{Macdonald_2012}.
+
 \bibliography{../manuscript/refs}

### 2024-09-27 proposal/project_proposal.tex
finished but could use some adjustments

@@
 idea behind the number is that it will show if a player has an overall positive or negative contribution when 
-they are playing. However, the plus/minus has a lot of drawbacks. For example, in ice hockey, players tend to play 
-on "lines", meaning the same three forwards and the same two defencemen tend to play at the same time. Because of this, 
-plus/minus is heavily influenced by who plays with who. Because of this, some argue that the plus/minus is more of a team 
-statistic rather than an individual statistic. Many of the criticisms of this statistic comes from the fact that
-there are a lot of confounding variables that influence how the plus/minus is calculated. According to Brian MacDonald 
+they are playing. However, the plus/minus has a lot of drawbacks. According to Brian MacDonald 
 "The traditional plus-minus statistic in hockey is highly dependent on a player’s teammates and opponents" \cite{Macdonald_2011}.
+In ice hockey, players tend to play on "lines", meaning the same three forwards and the same two defencemen tend 
+to play at the same time. Because of this, plus/minus is heavily influenced by who plays with who. Some argue that 
+the plus/minus is more of a team statistic rather than an individual statistic. Many of the criticisms of this statistic 
+comes from the fact that there are a lot of confounding variables that influence how the plus/minus is calculated.
 Alternatives to the plus/minus statistic have been developed, such as COSRI or Fenwick. Understanding the limitations and
 potential of plus/minus can help coaches, analysts, and fans better interpret player data. While advanced statistics are valuable, 
-plus/minus is still popular, especially among casual fans. The project explores whether the plus/minus still holds relevance or 
+plus/minus is still popular, especially among casual fans. This project explores whether the plus/minus still holds relevance or 
 whether it should be supplemented or replaced with other metrics.
@@
 \paragraph{Specific Aims}
-The goal of this project is to evaluate the effectiveness of the plus/minus statistic. Some of the information I am trying
-to determine is if the plus/minus statistic can be used to determine how good a player is both on offense and defense. I 
-also want to see if the plus/minus has any validity being used as an individual statistic or as a team statistic. Based on my 
-observations, I will compare the plus/minus statistic to other statistics that have a similar purpose such as CORSI or Fenwick. 
-These questions are critical as plus/minus is used in scouting and analysis, but its reliability has been debated. By comparing 
-it with modern metrics, I aim to shed light on its true value in player evaluation.
+The goal of this project is to evaluate the effectiveness of the plus/minus statistic. One of the topics of discussion is if the 
+plus/minus statistic can be used to determine how good a player is both on offense and defense. I also want to see if the plus/minus 
+has any validity being used as an individual statistic or as a team statistic. Based on my observations, I will compare the plus/minus 
+statistic to other statistics that have a similar purpose such as CORSI or Fenwick. These questions are critical as plus/minus is used 
+in scouting and analysis, but its reliability has been debated. By comparing it with modern metrics, I aim to shed light on its true 
+value in player evaluation.
 
@@
 will also compare plus/minus with other advanced hockey metrics, such as Corsi and Fenwick, to assess its overall utility.
-
 The first step of the project is to collect all relevant data. In this case, I will be exporting the data from all players
@@
 exist, since this will effect which models and variables I use. 
-
 In order to determine if or how plus/minus can be used to assess offensive and defensive contribution, I will use a ridge 
@@
 analysis. Similar work has been done in \cite{Macdonald_2012}. The coefficients from the ridge regression model will reveal 
-whether offensive or defensive metrics contribute more to explaining plus/minus. I plan on using a similar
+whether offensive or defensive metrics contribute more to explaining plus/minus. I plan on using a similar method for comparing
+the plus/minus with other advanced statistics, replacing the offesnive and defensive statistics with CORSI, Fenwick, and any other
+advanced statistics that are deemed relevant. When assessing if plus/minus is more influenced by individual performance or team 
+performance, I am considering using a mixed-effects ridge regression model. By using ridge regression with mixed-effects, it will 
+control for the potential influence of team quality on individual plus/minus values. This will help us determine how much of the 
+variability in plus/minus is explained by the team versus individual players.
 
 \paragraph{Discussion}
-What are the most challenge parts of the task?
-What are the limitations of your work? What is your fall-back plan if
-something unexpected happens?
-
-According to Brian MacDonald in his ridge regression analysis of the plus/minus, "Ridge regression frequently reduces the error 
-bounds in the estimates and improves the predictive performance of the model when collinearly exists in the data. Ridge regression 
-introduces bias in the estimates, but the tradeoff is typically worthwhile." \cite{Macdonald_2012}.
+The hardest part of this project will be finding ways to get around collinearity. Many of these statistics are highly correlated.
+The main goal to get around this is to implement ridge regression models. According to Brian MacDonald in his ridge regression analysis 
+of the plus/minus, "Ridge regression frequently reduces the error bounds in the estimates and improves the predictive performance 
+of the model when collinearly exists in the data. Ridge regression introduces bias in the estimates, but the tradeoff is typically 
+worthwhile." \cite{Macdonald_2012}. Mixed-effects models, principal component analysis, or other models could be used if ridge regression 
+proves to be ineffective. I am also assuming that the data that I am collecting from Natural Stat Trick is reliable and accurate. While 
+Natural Stat Trick provides advanced metrics, the quality and completeness of the data may vary. I will have to be extra dilligent in 
+the data cleaning process to ensure that variables of interest are available and consistent across the dataset.
 

### 2024-09-27 proposal/project_proposal.tex
spelling errors

@@
 comes from the fact that there are a lot of confounding variables that influence how the plus/minus is calculated.
-Alternatives to the plus/minus statistic have been developed, such as COSRI or Fenwick. Understanding the limitations and
+Alternatives to the plus/minus statistic have been developed, such as Corsi or Fenwick. Understanding the limitations and
 potential of plus/minus can help coaches, analysts, and fans better interpret player data. While advanced statistics are valuable, 
@@
 I collected data from the 2021-2022, 2022-2023, and 2023-2024 seasons. I am particularly interested in the advanced hockey 
-statistics that are collected for each player in the NHL. These statistics include CORSI, Fenwick, high-danger chances,
+statistics that are collected for each player in the NHL. These statistics include Corsi, Fenwick, high-danger chances,
 shooting percentages, and more. This data will also include more basic statistics such as points, goals, assists, plus/minus,
@@
 whether offensive or defensive metrics contribute more to explaining plus/minus. I plan on using a similar method for comparing
-the plus/minus with other advanced statistics, replacing the offesnive and defensive statistics with CORSI, Fenwick, and any other
+the plus/minus with other advanced statistics, replacing the offensive and defensive statistics with CORSI, Fenwick, and any other
 advanced statistics that are deemed relevant. When assessing if plus/minus is more influenced by individual performance or team 

### 2024-09-27 proposal/project_proposal.tex
changed some of the wording to make it easier to understand

@@
 \paragraph{Introduction}
-In hockey, a very common yet controversial statistic that is reported is the plus/minus statistic. Plus/minus is 
-calculated by taking all the even strength goals that were scored by their team while they were on the ice and 
-subtracting it by the total number of goals that were scored by the other team while they were on the ice. The 
-idea behind the number is that it will show if a player has an overall positive or negative contribution when 
-they are playing. However, the plus/minus has a lot of drawbacks. According to Brian MacDonald 
-"The traditional plus-minus statistic in hockey is highly dependent on a player’s teammates and opponents" \cite{Macdonald_2011}.
-In ice hockey, players tend to play on "lines", meaning the same three forwards and the same two defencemen tend 
-to play at the same time. Because of this, plus/minus is heavily influenced by who plays with who. Some argue that 
-the plus/minus is more of a team statistic rather than an individual statistic. Many of the criticisms of this statistic 
-comes from the fact that there are a lot of confounding variables that influence how the plus/minus is calculated.
-Alternatives to the plus/minus statistic have been developed, such as Corsi or Fenwick. Understanding the limitations and
+In ice hockey, a very common yet controversial statistic that is reported is the plus/minus statistic. It measures a player's impact 
+by calculating the difference between the even strength goals their team scores and concedes while they are on the ice. This metric
+aims to reflect whether a player contributes positively or negatively to the game However, the plus/minus has significant of drawbacks. 
+According to Brian MacDonald "The traditional plus-minus statistic in hockey is highly dependent on a player’s teammates and opponents" 
+\cite{Macdonald_2011}. Hockey players typically play in fixed "lines" where the same group of forwards and defensemen are on the ice 
+together. This makes plus/minus heavily influenced by a player's linemates rather than their individual performance. As a result, many
+argue that the statistic functions more as a reflection of team dynamics than as a precise indicator of individual contribution. 
+Furthermore, numerous confounding variables, such as the quality of the opponent and situational factors, make the calculation less 
+reliable. Alternatives to the plus/minus statistic have been developed, such as Corsi or Fenwick. Understanding the limitations and
 potential of plus/minus can help coaches, analysts, and fans better interpret player data. While advanced statistics are valuable, 
@@
 
-I am using data from the National Hockey League (NHL). The data is collected from the websites Natural Stat Trick and NHLEdge. 
-I collected data from the 2021-2022, 2022-2023, and 2023-2024 seasons. I am particularly interested in the advanced hockey 
-statistics that are collected for each player in the NHL. These statistics include Corsi, Fenwick, high-danger chances,
-shooting percentages, and more. This data will also include more basic statistics such as points, goals, assists, plus/minus,
-time on ice, etc.
+I am utilizing data from the National Hockey League (NHL), sourced from websites like Natural Stat Trick and NHLEdge.
+The data will be collected from the 2021-2022, 2022-2023, and 2023-2024 seasons. I am particularly interested in the advanced hockey 
+statistics that are collected for each player in the NHL. These advanced metrics include Corsi, Fenwick, high-danger chances, shooting 
+percentages, and other performance indicators. Additionally, the dataset features traditional statistics such as points, 
+goals, assists, plus/minus, and time on ice, providing a comprehensive view of player performance across multiple seasons.
 
@@
 The first step of the project is to collect all relevant data. In this case, I will be exporting the data from all players
-from the 2021- 2024 seasons into an Excel or csv file. Once the data has been collected, it needs to be cleaned. After the
-data has been prepared, I can start some initial analysis to see what variables are correlated and where issues of collinearity 
-exist, since this will effect which models and variables I use. 
-In order to determine if or how plus/minus can be used to assess offensive and defensive contribution, I will use a ridge 
-regression. Since the plus/minus statistic tends to deal with collinearity, a ridge regression might be more appropriate for 
-analysis. Similar work has been done in \cite{Macdonald_2012}. The coefficients from the ridge regression model will reveal 
+from the 2021- 2024 seasons into an Excel or csv file.  Following data collection, the next phase will involve data cleaning to 
+ensure accuracy and consistency. Once the dataset is prepared, initial analyses will identify correlations among variables 
+and pinpoint issues of collinearity, which could impact the selection of models and variables for further analysis.
+To evaluate how plus/minus can be employed to assess offensive and defensive contributions, I will implement ridge regression, 
+which is well-suited for handling collinearity concerns inherent in the plus/minus statistic, as 
+noted in prior research in \cite{Macdonald_2012}. The coefficients from the ridge regression model will reveal 
 whether offensive or defensive metrics contribute more to explaining plus/minus. I plan on using a similar method for comparing
@@
 \paragraph{Discussion}
-The hardest part of this project will be finding ways to get around collinearity. Many of these statistics are highly correlated.
-The main goal to get around this is to implement ridge regression models. According to Brian MacDonald in his ridge regression analysis 
+The most challenging aspect of this project will be addressing collinearity, as many of the statistics involved are highly correlated. 
+To mitigate this issue, I plan to implement ridge regression models. According to Brian MacDonald in his ridge regression analysis 
 of the plus/minus, "Ridge regression frequently reduces the error bounds in the estimates and improves the predictive performance 
 of the model when collinearly exists in the data. Ridge regression introduces bias in the estimates, but the tradeoff is typically 
-worthwhile." \cite{Macdonald_2012}. Mixed-effects models, principal component analysis, or other models could be used if ridge regression 
+worthwhile." \citet{Macdonald_2012}. Mixed-effects models, principal component analysis, or other models could be used if ridge regression 
 proves to be ineffective. I am also assuming that the data that I am collecting from Natural Stat Trick is reliable and accurate. While 
-Natural Stat Trick provides advanced metrics, the quality and completeness of the data may vary. I will have to be extra dilligent in 
-the data cleaning process to ensure that variables of interest are available and consistent across the dataset.
+Natural Stat Trick provides advanced metrics, the quality and completeness of the data may vary.  Therefore, I will be particularly diligent 
+in the data cleaning process to ensure that the variables of interest are available and consistent throughout the dataset. This 
+careful attention to detail will help enhance the resilience and reliability of the analysis and the validity of the results.
 

### 2024-10-10 code/data_cleaning.py
loaded the data

@@ -1,0 +1,33 @@
+"""
+Player - Player name.
+Team - Team or teams that the player has played for. Not displayed when filtering for specific teams.
+Position - Position or positions that the player has been listed as playing by the NHL.
+GP - Games played.
+TOI - Total amount of time played.
+CF - Count of Corsi for that player's team while that player is on the ice.
+CA - Count of Corsi against that player's team while that player is on the ice.
+CF% - Percentage of total Corsi while that player is on the ice that are for that player's team. CF*100/(CF+CA)
+Fenwick - any unblocked shot attempt (goals, shots on net and misses) outside of the shootout. Referred to as USAT by the NHL.
+FF - Count of Fenwick for that player's team while that player is on the ice.
+FA - Count of Fenwick against that player's team while that player is on the ice.
+FF% - Percentage of total Fenwick while that player is on the ice that are for that player's team. FF*100/(FF+FA)
+Shots - any shot attempt on net (goals and shots on net) outside of the shootout.
+SF - Count of Shots for that player's team while that player is on the ice.
+SA - Count of Shots against that player's team while that player is on the ice.
+SF% - Percentage of total Shots while that player is on the ice that are for that player's team. SF*100/(SF+SA)
+Goals - any goal, outside of the shootout.
+GF - Count of Goals for that player's team while that player is on the ice.
+GA - Count of Goals against that player's team while that player is on the ice.
+GF% - Percentage of total Goals while that player is on the ice that are for that player's team. GF*100/(GF+GA)
+Scoring Chances - a scoring chance, as originally defined by War-on-Ice
+"""
+
+import pandas as pd
+
+nhl_stats = pd.read_csv('data\Player Season Totals - Natural Stat Trick.csv')
+#print(nhl_stats.head())
+
+
+
+
+print(nhl_stats.dtypes)

### 2024-10-10 code/data_cleaning.py
loaded nhledge data

@@ -4,6 +4,7 @@ Team - Team or teams that the player has played for. Not displayed when filterin
 Position - Position or positions that the player has been listed as playing by the NHL.
 GP - Games played.
 TOI - Total amount of time played.
+Corsi - Any shot attempt (goals, shots on net, misses and blocks) outside of the shootout. Referred to as SAT by the NHL.
 CF - Count of Corsi for that player's team while that player is on the ice.
 CA - Count of Corsi against that player's team while that player is on the ice.
 CF% - Percentage of total Corsi while that player is on the ice that are for that player's team. CF*100/(CF+CA)
@@ -20,14 +21,72 @@ GF - Count of Goals for that player's team while that player is on the ice.
 GA - Count of Goals against that player's team while that player is on the ice.
 GF% - Percentage of total Goals while that player is on the ice that are for that player's team. GF*100/(GF+GA)
 Scoring Chances - a scoring chance, as originally defined by War-on-Ice
+SCF - Count of Scoring Chances for that player's team while that player is on the ice.
+SCA - Count of Scoring Chances against that player's team while that player is on the ice.
+SCF% - Percentage of total Scoring Chances while that player is on the ice that are for that player's team. SCF*100/(SCF+SCA)
+High Danger Scoring Chances - a scoring chance with a score of 3 or higher.
+HDCF - Count of High Danger Scoring Chances for that player's team while that player is on the ice.
+HDCA - Count of High Danger Scoring Chances against that player's team while that player is on the ice.
+HDCF% - Percentage of total High Danger Scoring Chances while that player is on the ice that are for that player's team. HDCF*100/(HDCF+HDCA)
+High Danger Goals - goals generated from High Danger Scoring Chances
+HDGF - Count of Goals off of High Danger Scoring Chances for that player's team while that player is on the ice.
+HDGA - Count of Goals off of High Danger Scoring Chances against that player's team while that player is on the ice.
+HDGF% - Percentage of High Danger Goals while that player is on the ice that are for that player's team. HDGF*100/(HDGF+HDGA)
+Medium Danger Scoring Chances - a scoring chance with a score of exactly 2.
[…]

### 2024-10-10 code/data_cleaning.py
fixed loading problem and working on fixing duplicate columns

@@ -70,23 +70,46 @@ from functools import reduce
 nst = pd.read_csv('data\Player Season Totals - Natural Stat Trick.csv')
 #print(nst.head())
 
-edge1 = pd.read_excel(f'data\Summary(1).xlsx')
-edge2 = pd.read_excel(f'data\Summary(2).xlsx')
-edge3 = pd.read_excel(f'data\Summary(3).xlsx')
-edge4 = pd.read_excel(f'data\Summary(4).xlsx')
-edge5 = pd.read_excel(f'data\Summary(5).xlsx')
-edge6 = pd.read_excel(f'data\Summary(6).xlsx')
-edge7 = pd.read_excel(f'data\Summary(7).xlsx')
-edge8 = pd.read_excel(f'data\Summary(8).xlsx')
-edge9 = pd.read_excel(f'data\Summary(9).xlsx')
-edge10 = pd.read_excel(f'data\Summary(10).xlsx')
-edge11 = pd.read_excel(f'data\Summary(11).xlsx')
-edge12 = pd.read_excel(f'data\Summary(12).xlsx')
-edge13 = pd.read_excel(f'data\Summary(13).xlsx')
+edge1 = pd.read_excel('data\\Summary (1).xlsx')
+edge2 = pd.read_excel('data\\Summary (2).xlsx')
+edge3 = pd.read_excel('data\\Summary (3).xlsx')
+edge4 = pd.read_excel('data\Summary (4).xlsx')
+edge5 = pd.read_excel('data\Summary (5).xlsx')
+edge6 = pd.read_excel('data\Summary (6).xlsx')
+edge7 = pd.read_excel('data\Summary (7).xlsx')
+edge8 = pd.read_excel('data\Summary (8).xlsx')
+edge9 = pd.read_excel('data\Summary (9).xlsx')
+edge10 = pd.read_excel('data\Summary (10).xlsx')
+edge11 = pd.read_excel('data\Summary (11).xlsx')
+edge12 = pd.read_excel('data\Summary (12).xlsx')
+edge13 = pd.read_excel('data\Summary (13).xlsx')
 
 edges = [edge1, edge2, edge3, edge4, edge5, edge6, edge7, edge8, edge9, edge10, edge11, edge12, edge13]
 
-nhl_edge = reduce(lambda left, right: pd.merge(left, right, on='team', how='outer'), edges)
+import pandas as pd
+from functools import reduce
+
+# Custom function to merge only if values are the same for duplicate columns
[…]

### 2024-10-11 code/data_cleaning.py
attempting to resolve issue with merging columns

@@ -89,27 +89,42 @@ edges = [edge1, edge2, edge3, edge4, edge5, edge6, edge7, edge8, edge9, edge10, 
 import pandas as pd
 from functools import reduce
 
-# Custom function to merge only if values are the same for duplicate columns
-def merge_duplicates(df, suffixes=('_x', '_y')):
-    # Identify columns with suffixes
-    cols_x = [col for col in df.columns if col.endswith(suffixes[0])]
-    cols_y = [col for col in df.columns if col.endswith(suffixes[1])]
+import pandas as pd
+from functools import reduce
+
+import pandas as pd
+from functools import reduce
+
+# Helper function to resolve conflicts between two DataFrames during the merge
+def merge_and_resolve_conflicts(left, right, on='Player', suffixes=('_x', '_y')):
+    merged_df = pd.merge(left, right, on=on, how='outer', suffixes=suffixes)
     
-    # Iterate over columns with suffixes
+    # Find columns with suffixes
+    cols_x = [col for col in merged_df.columns if col.endswith(suffixes[0])]
+    cols_y = [col for col in merged_df.columns if col.endswith(suffixes[1])]
+    
     for col_x in cols_x:
+        # Get the corresponding _y column name
         col_y = col_x.replace(suffixes[0], suffixes[1])
-        if col_y in cols_y:
-            # Check if values are equal in both columns
-            df[col_x.replace(suffixes[0], '')] = df[col_x].combine_first(df[col_y])
-            # Drop the duplicate columns
-            df.drop([col_x, col_y], axis=1, inplace=True)
-    return df
+        
+        if col_y in merged_df.columns:
+            # Create the original column name (without suffixes)
+            col_base = col_x.replace(suffixes[0], '')
+            
+            # Use combine_first() to keep non-null values from both columns
+            merged_df[col_base] = merged_df[col_x].combine_first(merged_df[col_y])
+            
+            # Drop the suffixed columns
+            merged_df.drop([col_x, col_y], axis=1, inplace=True)
+    
+    return merged_df
 
-# Perform the merge and resolve duplicates
-nhl_edge = reduce(lambda left, right: pd.merge(left, right, on='Player', how='outer', suffixes=('_x', '_y')), edges)
+# Merging multiple DataFrames using reduce and custom conflict resolution
+nhl_edge = reduce(lambda left, right: merge_and_resolve_conflicts(left, right), edges)
 
[…]

### 2024-10-11 code/data_cleaning.py
dropped some columns and filtered GP

@@ -122,9 +122,14 @@ def merge_and_resolve_conflicts(left, right, on='Player', suffixes=('_x', '_y'))
 # Merging multiple DataFrames using reduce and custom conflict resolution
 nhl_edge = reduce(lambda left, right: merge_and_resolve_conflicts(left, right), edges)
 
-# Check the result
-print(nhl_edge.head())
+nhl_edge = nhl_edge.drop('FOW%', axis=1)
 
+nhl_edge = nhl_edge[nhl_edge['GP'] > 15]
+#print(nhl_edge.head())
+print(nhl_edge.dtypes)
 
-print(nst.dtypes)
+#To Do:
+# remove players with less than 15 games played
+# fix format of TOI/GP
+# change S and S% to float
 

### 2024-10-14 code/data_cleaning.py
fixed the format of TOI/GP

@@ -89,12 +89,6 @@ edges = [edge1, edge2, edge3, edge4, edge5, edge6, edge7, edge8, edge9, edge10, 
 import pandas as pd
 from functools import reduce
 
-import pandas as pd
-from functools import reduce
-
-import pandas as pd
-from functools import reduce
-
 # Helper function to resolve conflicts between two DataFrames during the merge
 def merge_and_resolve_conflicts(left, right, on='Player', suffixes=('_x', '_y')):
     merged_df = pd.merge(left, right, on=on, how='outer', suffixes=suffixes)
@@ -122,14 +116,28 @@ def merge_and_resolve_conflicts(left, right, on='Player', suffixes=('_x', '_y'))
 # Merging multiple DataFrames using reduce and custom conflict resolution
 nhl_edge = reduce(lambda left, right: merge_and_resolve_conflicts(left, right), edges)
 
-nhl_edge = nhl_edge.drop('FOW%', axis=1)
+nhl_edge = nhl_edge.drop(['FOW%', 'PIM'], axis=1)
 
 nhl_edge = nhl_edge[nhl_edge['GP'] > 15]
-#print(nhl_edge.head())
+
+
+nhl_edge.fillna(0, inplace=True)
+
+print(nhl_edge.head())
 print(nhl_edge.dtypes)
 
+# Function to convert 'Minutes:Seconds' to float
+def convert_to_float(time_str):
+    minutes, seconds = map(int, time_str.split(':'))
+    return minutes + (seconds / 60.0)
+
+# Apply the function to the column
+nhl_edge['TOI/GP'] = nhl_edge['TOI/GP'].apply(convert_to_float)
+
+nhl_edge.fillna(0, inplace=True)
+# View the result
+print(nhl_edge.head(-5))
+print(nhl_edge.dtypes)
 #To Do:
-# remove players with less than 15 games played
-# fix format of TOI/GP
 # change S and S% to float
 

### 2024-10-30 code/analysis.py
did the offensive vs defensive contribution

@@ -24,10 +24,97 @@ plt.show()
 plus_minus_corr = nhl.select_dtypes(include=['number']).corr()['+/-']
 print(plus_minus_corr)
 
-# how plus/minus can be employed to assess offensive and defensive contribution with ridge regression
+print(nhl.columns)
+# how plus/minus can be employed to assess offensive and defensive contribution with ridge regression/regular regression
+
+import pandas as pd
+from sklearn.linear_model import Ridge
+from sklearn.model_selection import train_test_split, cross_val_score
+from sklearn.preprocessing import StandardScaler
+
+# Define target and features (Offensive vs. Defensive)
+# Offensive variables: CF, FF, SF, GF, SCF, HDCF, HDGF, GF%, G, A, P/GP, CF%, SCF%, On-Ice SH%
+# Defensive variables: CA, FA, SA, GA, SCF, HDCA, HDGA, On-Ice SV%
+X = nhl[['GF%', 'SCF%', 'HDCF%', 'PDO', 'On-Ice SH%', 'On-Ice SV%']]
+y = nhl['+/-']  # Plus/minus as the target variable
+
+# Scale the features for Ridge Regression
+scaler = StandardScaler()
+X_scaled = scaler.fit_transform(X)
+
+# Split the data into training and test sets
+X_train, X_test, y_train, y_test = train_test_split(X_scaled, y, test_size=0.2, random_state=42)
+
+# Fit Ridge Regression Model
+ridge = Ridge(alpha=1.0)  # alpha is the regularization strength (lambda in ridge regression)
+ridge.fit(X_train, y_train)
+
+# Evaluate the model using cross-validation
+cv_scores = cross_val_score(ridge, X_train, y_train, cv=5)
+print("Cross-Validation Scores (Offensive/Defensive):", cv_scores)
+
+# Predict on the test set
+y_pred = ridge.predict(X_test)
+
+
+# Offensive alone
+X_o = nhl[['CF', 'FF', 'SF', 'GF', 'SCF', 'HDCF', 'HDGF', 'GF%', 'G', 'A', 'P/GP', 'CF%', 'SCF%', 'On-Ice SH%']]
+y_o = nhl['+/-']  # Plus/minus as the target variable
+
+# Scale the features for Ridge Regression
+scaler = StandardScaler()
+X_o_scaled = scaler.fit_transform(X_o)
+
+# Split the data into training and test sets
[…]

### 2024-10-31 code/data_cleaning.py
fixed null values in nst

@@ -68,7 +68,19 @@ Off. Zone Faceoff % - Percentage of faceoffs in the offensive zone for which the
 import pandas as pd
 from functools import reduce
 nst = pd.read_csv('data\Player Season Totals - Natural Stat Trick.csv')
-#print(nst.head())
+print(nst.head())
+
+nst['SF%'].fillna(0, inplace=True)
+nst['GF%'].fillna(0, inplace=True)
+nst['SCF%'].fillna(0, inplace=True)
+nst['HDCF%'].fillna(0, inplace=True)
+nst['HDGF%'].fillna(0, inplace = True)
+nst['On-Ice SH%'].fillna(0, inplace=True)
+nst['On-Ice SV%'].fillna(0, inplace=True)
+nst['PDO'].fillna(0, inplace=True)
+nst['Off. Zone Start %'].fillna(0, inplace=True)
+nst['Off. Zone Faceoff %'].fillna(0, inplace=True)
+print(nst.dtypes)
 
 edge1 = pd.read_excel('data\\Summary (1).xlsx')
 edge2 = pd.read_excel('data\\Summary (2).xlsx')
@@ -135,13 +147,10 @@ def convert_to_float(time_str):
 nhl_edge['TOI/GP'] = nhl_edge['TOI/GP'].apply(convert_to_float)
 
 nhl_edge.fillna(0, inplace=True)
-#To Do:
-# change S and S% to float
+
 def clean_columns(df, columns):
     for col in columns:
-        # Replace dashes with NaN (or 0 if you prefer)
         df[col] = df[col].replace('--', 0)
-        # Convert to numeric type
         df[col] = pd.to_numeric(df[col], errors='coerce')
     return df
 
@@ -159,4 +168,7 @@ print(nhl_edge[['Player', 'S', 'S%']].head())
 csv_file_path = 'data/cleaned_nhl_edge.csv'
 
 # Save the nhl_edge DataFrame as a CSV file
-nhl_edge.to_csv(csv_file_path, index=False)
+#nhl_edge.to_csv(csv_file_path, index=False)
+nst = clean_columns(nst, ['SF%', 'GF%', 'SCF%', 'HDCF%','HDGF%', 'On-Ice SH%', 'On-Ice SV%', 'PDO', 'Off. Zone Start %', 'Off. Zone Faceoff %'])
+print(nst.dtypes)
+nst.to_csv('data/cleaned_nst.csv', index=False)

### 2024-10-31 code/analysis.py
did the analysis

@@ -1,7 +1,7 @@
 import pandas as pd
 
 nhl_edge = pd.read_csv('data\cleaned_nhl_edge.csv')
-nst = pd.read_csv('data\Player Season Totals - Natural Stat Trick.csv')
+nst = pd.read_csv('data\cleaned_nst.csv')
 nhl = pd.merge(nhl_edge, nst, on='Player', how='inner')  # 'inner' merge keeps only players present in both datasets
 
 # Identify collinearity
@@ -91,11 +91,11 @@ X_scaled_d = scaler.fit_transform(X_d)
 X_train_d, X_test_d, y_train_d, y_test_d = train_test_split(X_scaled_d, y_d, test_size=0.2, random_state=42)
 
 # Fit Ridge Regression Model
-ridge = Ridge(alpha=1.0)  # alpha is the regularization strength (lambda in ridge regression)
-ridge.fit(X_train_d, y_train_d)
+ridge_d = Ridge(alpha=1.0)  # alpha is the regularization strength (lambda in ridge regression)
+ridge_d.fit(X_train_d, y_train_d)
 
 # Evaluate the model using cross-validation
-cv_scores_d = cross_val_score(ridge, X_train_d, y_train_d, cv=5)
+cv_scores_d = cross_val_score(ridge_d, X_train_d, y_train_d, cv=5)
 print("Cross-Validation Scores (Defensive):", cv_scores_d)
 
 
@@ -109,7 +109,7 @@ nhl['TeamMean_GF%'] = nhl.groupby('Team')['GF%'].transform('mean')
 
 # Define features (team and player contributions)
 X = nhl[['CF%', 'GF%', 'TeamMean_CF%', 'TeamMean_GF%']]
-y = nhl['PlusMinus']
+y = nhl['+/-']
 
 # Fit a Mixed-Effects model with team as a random effect
 md = MixedLM(y, X, groups=nhl['Team'])
@@ -118,3 +118,47 @@ print(mdf.summary())
 
 
 # Plus minus vs other evaluators
+
+X_corsi = nhl[['GF%', 'SCF%', 'HDCF%', 'PDO', 'On-Ice SH%', 'On-Ice SV%', 'G', 'A']]
+y_cf = nhl['CF%']  # Plus/minus as the target variable
+
+# Scale the features for Ridge Regression
+scaler = StandardScaler()
+X_scaled_c = scaler.fit_transform(X_corsi)
+
+# Split the data into training and test sets
[…]

### 2024-11-01 code/analysis.py
filtered out players with less than 25 games played

@@ -3,7 +3,7 @@ import pandas as pd
 nhl_edge = pd.read_csv('data\cleaned_nhl_edge.csv')
 nst = pd.read_csv('data\cleaned_nst.csv')
 nhl = pd.merge(nhl_edge, nst, on='Player', how='inner')  # 'inner' merge keeps only players present in both datasets
-
+nhl = nhl[nhl['GP_x'] > 25]
 # Identify collinearity
 #print(nhl_edge.describe())
 # Select only numerical columns

### 2024-11-01 code/analysis.py
generated the correlation matrix

@@ -10,6 +10,9 @@ nhl = nhl[nhl['GP_x'] > 25]
 nhl_edge_numeric = nhl_edge.select_dtypes(include=['number'])
 correlation_matrix = nhl_edge_numeric.corr()
 #print(correlation_matrix)
+with open("correlation_matrix.tex", "w") as file:
+    file.write(correlation_matrix.to_latex())
+
 
 import matplotlib.pyplot as plt
 """

### 2024-11-01 manuscript/statspaper.tex
worked on methods and started results

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

%% for double spacing
\usepackage{setspace}

% for space filling
\usepackage{lipsum}
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

% Get line numbers for ease of referencing for the reviewers
\usepackage[]{lineno}
\linenumbers*[1]
% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
        \expandafter\let\csname old#1\expandafter\endcsname\csname 
        #1\endcsname
        \expandafter\let\csname oldend#1\expandafter\endcsname\csname 
        end#1\endcsname
        \renewenvironment{#1}%
        {\linenomath\csname old#1\endcsname}%
        {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
        \patchAmsMathEnvironmentForLineno{#1}%
        \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
        \patchBothAmsMathEnvironmentsForLineno{equation}%
        \patchBothAmsMathEnvironmentsForLineno{align}%
        \patchBothAmsMathEnvironmentsForLineno{flalign}%
        \patchBothAmsMathEnvironmentsForLineno{alignat}%
        \patchBothAmsMathEnvironmentsForLineno{gather}%
        \patchBothAmsMathEnvironmentsForLineno{multline}%
}

%% meta data

\title{Hockey's Most Controversial Statistic: An Analysis of the Effectiveness of the Plus-Minus Statistic}
\author{Melanie Desroches\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\begin{abstract}
Here is the abstract.  
\end{abstract}

\doublespacing

\section{Introduction}
\label{sec:intro}

Use this section to answer three questions:
Why is the topic important/interesting?
What has been done on this topic in the literature?
What is your contribution?

Within the realm of sports, many different statistics are used to determine how good a player is. This can include points,
assists, time played, etc. But is there just one valaue that can be used to determine the effectiveness of a player? This is an 
important topic, as fans, coaches, players, and general managers often want to get a better understanding of how an individual is
impacting their team. It is one thing to watch the players to determine how good a player is, often refered to as the "eye test" but 
with so many players, it is hard to quantify this. Enter the plus-minus statistic. Plus-minus is generally calculated by adding all the points
scored by their team while they were playing and subtracting by points scored by the opposition while they were playing. The idea is that 
if a player has a generally positive impact on their team, they will have a highly positive plus-minus. The two main sports where this
statistic is used is basketball and ice hockey.

While the plus-minus statistic is a great idea in theory, it does not come without it's weaknesses. The biggest drawback to this
statistic is its issues with independece. Take ice hockey for example. Many ice hockey players tend to play on "lines", meaning the same three 
forwards tend to play together and the same two defensemen usually play together. As a result, the performance of one player is highly dependent 
on the performance of their linemates. In "A Regression-based Adjusted Plus-Minus Statistic for NHL Players" \cite{Macdonald_2011}, Brian
MacDonald provides a perfect example of this with the Henrik and Daniel Sedin, Swedish twins who played together for the Vancouver Canucks. 
 \cite[Daniel spent 92\% of his playing time with Henrik, the highest percentage of any other player combination where both players have
played over 700 minutes. Because of this high colinearity between the twins, it is difficult to separate the individual effect that each 
player has on the net goals scored on the ice.]{Macdonald_2012} Many critics of the plus-minus argue that plus-minus is more of a team 
statistic, since it is heavily influenced by team dynamics rather than individual contributions. Furthermore, numerous confounding variables, 
such as the quality of the opponent and situational factors, make the calculation less reliable. As a result of these criticisms, alternatives 
to the plus-minus statistic have been developed, such as Corsi or Fenwick, which is calculated using shot attempts rather than goals.

The goal of this project will be to shed light on this controversial statistic and determine if plus-minus is truely an effective predictor
to individual contribution, specifically in the National Hockey League (hereforth refered to as the NHL). Is the plus-minus a good evaluator 
of offensive and defensive output? Is the plus-minus more reflective of how well a player is doing than how well a player is preforming? 
Are there better alternatives to evaluating individual player contribution? These are the questions that will be answered from this analysis.

% roadmap
The rest of the paper is organized as follows.
The data will be presented in Section~\ref{sec:data}.
The methods are described in Section~\ref{sec:meth}.
The results are reported in Section~\ref{sec:resu}.
A discussion concludes in Section~\ref{sec:disc}.

\section{Data}
\label{sec:data}

Use this section to describe the data that helps to answer your research
questions.

The data used to perform this analysis was collected from Natural Stat Trick and the NHL website. The observations in the data
is from all players that played in the National Hockey League from the 2021-2022, 2022-2023, and 2023-2024 seasons. Only players
that have played in over 25 games were used in the analysis. This is ensure that the players being evaluated were all of similar 
calliber and were regular NHL players. The columns of the dataset will be combined from the two data sources. The following is a 
description of each of the columns in the dataset obtained from the Natural Stat Trick website:

Player - Player name.

Team - Team or teams that the player has played for. Not displayed when filtering for specific teams.

Position - Position or positions that the player has been listed as playing by the NHL.

GP - Games played.

TOI - Total amount of time played.

Corsi - Any shot attempt (goals, shots on net, misses and blocks) outside of the shootout. Referred to as SAT by the NHL.

CF - Count of Corsi for that player's team while that player is on the ice.

CA - Count of Corsi against that player's team while that player is on the ice.

CF\% - Percentage of total Corsi while that player is on the ice that are for that player's team. CF*100/(CF+CA)

Fenwick - any unblocked shot attempt (goals, shots on net and misses) outside of the shootout. Referred to as USAT by the NHL.

FF - Count of Fenwick for that player's team while that player is on the ice.

FA - Count of Fenwick against that player's team while that player is on the ice.

FF\% - Percentage of total Fenwick while that player is on the ice that are for that player's team. FF*100/(FF+FA)

Shots - any shot attempt on net (goals and shots on net) outside of the shootout.

SF - Count of Shots for that player's team while that player is on the ice.

SA - Count of Shots against that player's team while that player is on the ice.

SF\% - Percentage of total Shots while that player is on the ice that are for that player's team. SF*100/(SF+SA)

Goals - any goal, outside of the shootout.

GF - Count of Goals for that player's team while that player is on the ice.

GA - Count of Goals against that player's team while that player is on the ice.

GF\% - Percentage of total Goals while that player is on the ice that are for that player's team. GF*100/(GF+GA)

Scoring Chances - a scoring chance, as originally defined by War-on-Ice

SCF - Count of Scoring Chances for that player's team while that player is on the ice.

SCA - Count of Scoring Chances against that player's team while that player is on the ice.

SCF\% - Percentage of total Scoring Chances while that player is on the ice that are for that player's team. SCF*100/(SCF+SCA)

High Danger Scoring Chances - a scoring chance with a score of 3 or higher.

HDCF - Count of High Danger Scoring Chances for that player's team while that player is on the ice.

HDCA - Count of High Danger Scoring Chances against that player's team while that player is on the ice.

HDCF\% - Percentage of total High Danger Scoring Chances while that player is on the ice that are for that player's team. HDCF*100/(HDCF+HDCA)

High Danger Goals - goals generated from High Danger Scoring Chances

HDGF - Count of Goals off of High Danger Scoring Chances for that player's team while that player is on the ice.

HDGA - Count of Goals off of High Danger Scoring Chances against that player's team while that player is on the ice.

HDGF\% - Percentage of High Danger Goals while that player is on the ice that are for that player's team. HDGF*100/(HDGF+HDGA)

Medium Danger Scoring Chances - a scoring chance with a score of exactly 2.

MDCF - Count of Medium Danger Scoring Chances for that player's team while that player is on the ice.

MDCA - Count of Medium Danger Scoring Chances against that player's team while that player is on the ice.

MDCF\% - Percentage of total Medium Danger Scoring Chances while that player is on the ice that are for that player's team. MDCF*100/(MDCF+MDCA)

Medium Danger Goals - goals generated from Medium Danger Scoring Chances

MDGF - Count of Goals off of Medium Danger Scoring Chances for that player's team while that player is on the ice.

MDGA - Count of Goals off of Medium Danger Scoring Chances against that player's team while that player is on the ice.

MDGF\% - Percentage of Medium Danger Goals while that player is on the ice that are for that player's team. MDGF*100/(MDGF+MDGA)

Low Danger Scoring Chances - a scoring chance with a score of 1 or less. Does not include any attempts from the attacking team's neutral or defensive zone.

LDCF - Count of Low Danger Scoring Chances for that player's team while that player is on the ice.

LDCA - Count of Low Danger Scoring Chances against that player's team while that player is on the ice.

LDCF\% - Percentage of total Low Danger Scoring Chances while that player is on the ice that are for that player's team. LDCF*100/(LDCF+LDCA)

Low Danger Goals - goals generated from Low Danger Scoring Chances

LDGF - Count of Goals off of Low Danger Scoring Chances for that player's team while that player is on the ice.

LDGA - Count of Goals off of Low Danger Scoring Chances against that player's team while that player is on the ice.

LDGF\% - Percentage of Low Danger Goals while that player is on the ice that are for that player's team. LDGF*100/(LDGF+LDGA)

PDO

SH\% - Percentage of Shots for that player's team while that player is on the ice that were Goals. GF*100/SF

SV\% - Percentage of Shots against that player's team while that player is on the ice that were not Goals. GA*100/SA

PDO - Shooting percentage plus save percentage. (GF/SF)+(GA/SA)

Starts

Off. Zone Starts - Number of shifts for the player that started with an offensive zone faceoff.

Neu. Zone Starts - Number of shifts for the player that started with an neutral zone faceoff.

Def. Zone Starts - Number of shifts for the player that started with an defensive zone faceoff.

On The Fly Starts - Number of shifts for the player that started during play (without a faceoff).

Off. Zone Start \% - Percentage of starts for the player that were Offensive Zone Starts, excluding Neutral Zone and On The Fly Starts. Off. Zone Starts*100/(Off. Zone Starts+Def. Zone Starts)

Faceoffs

Off. Zone Faceoffs - Number of faceoffs in the offensive zone for which the player was on the ice.

Neu. Zone Faceoffs - Number of faceoffs in the neutral zone for which the player was on the ice.

Def. Zone Faceoffs - Number of faceoffs in the defensive zone for which the player was on the ice.

Off. Zone Faceoff \% - Percentage of faceoffs in the offensive zone for which the player was on the ice, excluding neutral zone faceoffs. Off. Zone Faceoffs*100/(Off. Zone Faceoffs+Def. Zone Faceoffs)

The goal of this dataset was to provide a broader scope of variables that evaluate a players performance. By introducing more
advanced and uncovential statistics, there are more factors to consider in evaluating a player's effectiveness on the ice.

\section{Methods}
\label{sec:meth}

Use this section to present the methodologies that will generate results by
analyzing the data.

The main issue with the data is the violation of independence. This can be seen by creating a confusion matrix of the some 
of the more basic nhl statistics.

\begin{table}[h!]
  \centering
  \caption{Correlation Matrix of Variables}
  \input{correlation_matrix.tex}
  \label{tab:correlation_matrix}
\end{table}

There is a high correlation between different shot related statistics and point related statistics. For example, there is a 
0.826371 correlation between Goals and Shots. This makes since because in order to score a goal, the player needs to shoot first.
Correlation analysis can be used to identify the strength and direction of the relationship between various advanced metrics and 
the plus/minus statistic. This was done by calculating the correlation coefficient between +/- and each metric (like CF\%, GF\%, SCF\%, etc.).
The correlation values, ranging from -1 to 1, tell us how closely each metric aligns with the plus/minus. Higher positive correlations 
(e.g., with GF\% and PDO) indicate metrics that vary similarly to plus/minus, suggesting that they may capture overlapping aspects of 
offensive or defensive performance.

Due to the presence of correlation among variables, ridge regression was performed in order to identify how plus/minus can be employed 
to assess offensive and defensive contribution. It was also used to determine if other advanced metrics Corsi or Fenwick have the same, 
better, or worse predictive abilities compared to plus-minus. Ridge regression, a form of linear regression, is valuable when dealing 
with datasets with multicollinearity, where predictor variables have some sort of relationship with one another. This applies to hockey metrics 
like Corsi, Fenwick, scoring chances, etc. since these stats often interact or overlap in measuring aspects of performance.

Coupled with ridge regression, cross validation was utilized to assess the predictive power of different groups of variables (offensive, 
defensive, possession-based) on the plus/minus statistic. Cross-validation splits the data into training and testing sets multiple times, 
computing a model's predictive accuracy each time. In this case, the data was split into five subsets. In each iteration, one subset was
kept as the test set and the other four were used to train the ridge regression model. The R-sqared value was reported for each iteration
to measure how well offensive, defensive, and possession-based metrics (like Corsi and Fenwick) predict plus/minus.

In order to separate individual contributions to plus/minus from team-level effects, a mixed-effects model was used. Mixed-effects models 
allow for random effects, capturing variability at the team level, while fixed effects capture individual-level metrics. Individual metrics 
like CF\%, GF\%, and others were included that directly measure a player's on-ice performance. The model considered team-level averages 
(e.g., TeamMeanGF\%), accounting for how a player’s team context might influence their plus/minus.

\section{Results}
\label{sec:resu}

Correlation with Plus/Minus (+/-):

Metrics such as GF\%, SF\%, PDO, xGF\%, and HDGF\% show a relatively strong correlation with plus/minus, which aligns with 
expectations since these metrics also track goal-related contributions.

Cross-Validation Scores:

Offensive/Defensive Combined: Averages around 0.53, indicating that these metrics explain a moderate proportion of the variance in plus/minus.
Offensive Only: Cross-validation scores are around 0.63, suggesting that offensive metrics alone are somewhat predictive of plus/minus, 
reflecting the offensive impact.
Defensive Only: With higher scores around 0.71, defensive metrics seem to have a greater influence on the plus/minus, which aligns with 
the role of plus/minus in capturing defensive contributions.
Corsi and Fenwick: Both show very high cross-validation scores, with Fenwick performing slightly better, indicating that these 
possession-based metrics are strong predictors.

Mixed Effects Model:

The significant coefficients for both GF\% and team averages, negative for TeamMeanCF\% and positive for TeamMeanGF\%, imply that 
team-level factors substantially impact individual plus/minus scores.
This aligns with the idea that plus/minus is both a team and individual statistic, as team-level scoring factors are significant predictors.

\begin{table}[tbp]
  \caption{Cross Validation Scores}
  \label{tab:rv}
\centering
\begin{tabular}{rrr}
  \toprule
Both & Offensive & Defensive \\ 
  \midrule
  0.56777061 & 0.64692431 & 0.73265605 \\ 
  0.51687324 & 0.62667921 & 0.70862251 \\ 
  0.49576518 & 0.62263538 & 0.69467537 \\ 
  0.49576518 & 0.58484799 & 0.65315835 \\ 
  0.61054685 & 0.69145036 & 0.76239753 \\ 
   \bottomrule
\end{tabular}
\end{table}

\section{Discussion}
\label{sec:disc}

What are the main contributions again?

What are the limitations of this study?

What are worth pursuing further in the future?

\bibliography{refs}
\bibliographystyle{mcap}

\end{document}

### 2024-11-01 manuscript/statspaper.tex
added results

@@
 
-The main issue with the data is the violation of independence. This can be seen by creating a confusion matrix of the some 
-of the more basic nhl statistics.
+\subsection{Correlation Analysis}
+The main issue with the data is multicollinearity, where independent variables are highly correlated. A correlation matrix was created 
+with some of the more basic nhl statistics to highlight this issue, shown in Table \ref{tab:correlation_matrix}.
 
@@
 Correlation analysis can be used to identify the strength and direction of the relationship between various advanced metrics and 
-the plus/minus statistic. This was done by calculating the correlation coefficient between +/- and each metric (like CF\%, GF\%, SCF\%, etc.).
-The correlation values, ranging from -1 to 1, tell us how closely each metric aligns with the plus/minus. Higher positive correlations 
-(e.g., with GF\% and PDO) indicate metrics that vary similarly to plus/minus, suggesting that they may capture overlapping aspects of 
-offensive or defensive performance.
+the plus/minus statistic. Metrics with high positive correlations (e.g., GF\% and PDO) suggest a strong alignment with the plus/minus, 
+indicating that they may reflect similar aspects of offensive or defensive performance. This was done by calculating the correlation 
+coefficient between +/- and each metric (like CF\%, GF\%, SCF\%, etc.).The correlation values, ranging from -1 to 1, tell us how closely 
+each metric aligns with the plus/minus. Higher positive correlations (e.g., with GF\% and PDO) indicate metrics that vary similarly to 
+plus/minus, suggesting that they may capture overlapping aspects of offensive or defensive performance.
+
+\subsection{Ridge Regression}
 
@@
 to assess offensive and defensive contribution. It was also used to determine if other advanced metrics Corsi or Fenwick have the same, 
-better, or worse predictive abilities compared to plus-minus. Ridge regression, a form of linear regression, is valuable when dealing 
-with datasets with multicollinearity, where predictor variables have some sort of relationship with one another. This applies to hockey metrics 
-like Corsi, Fenwick, scoring chances, etc. since these stats often interact or overlap in measuring aspects of performance.
+better, or worse predictive abilities compared to plus-minus. Ridge regression, a form of regularized linear regression, is beneficial 
+in handling datasets where predictor variables are highly interrelated, as is the case with advanced hockey metrics. Ridge regression 
+estimates the contribution of metrics like Corsi, Fenwick, and scoring chances while controlling for their interdependencies. This 
+applies to hockey metrics like Corsi, Fenwick, scoring chances, etc. since these statistics often interact or overlap in measuring aspects 
+of performance.
 
+\subsection{Cross-Validation}
 Coupled with ridge regression, cross validation was utilized to assess the predictive power of different groups of variables (offensive, 
@@
 kept as the test set and the other four were used to train the ridge regression model. The R-sqared value was reported for each iteration
-to measure how well offensive, defensive, and possession-based metrics (like Corsi and Fenwick) predict plus/minus.
+to measure how well offensive, defensive, and possession-based metrics (like Corsi and Fenwick) predict plus/minus. This process ensures 
+that the model’s performance is consistent across different data partitions.
 
+\subsection{Mixed Model Effects}
 In order to separate individual contributions to plus/minus from team-level effects, a mixed-effects model was used. Mixed-effects models 
@@
 
-Correlation with Plus/Minus (+/-):
+The results of taking the correlation coefficients of plus-minus versus every other column is as follows:
 
-Metrics such as GF\%, SF\%, PDO, xGF\%, and HDGF\% show a relatively strong correlation with plus/minus, which aligns with 
-expectations since these metrics also track goal-related contributions.
+\begin{table}[tbp]
+  \caption{Cross Validation Scores}
+  \label{tab:rv}
+\centering
+\begin{tabular}{rrr}
+  \toprule
+Variable & Correlation Coefficient \\ 
+  \midrule
+  GP & 0.232144
+  G & 0.237058
+  A & 0.342538
+  P & 0.314842
+  +/- & 1.000000
+  P/GP & 0.277666
+  EVG & 0.239076
+  EVP & 0.336372
+  PPG & 0.187284
+  PPP & 0.226803
+  SHG & 0.177479
+  SHP & 0.282570
+  OTG & 0.161693
+  GWG & 0.305998
+  S & 0.255841
+  S\% & 0.048282
+  TOI/GP & 0.253234
+  TOI & 0.291711
+  CF & 0.369200
+  CA & 0.210752
+  CF\% & 0.533927
+  FF & 0.373707
+  FA & 0.213121
+  FF\% & 0.553993
+  SF & 0.376144
+  SA & 0.211659
+  SF\% & 0.568873
+  GF & 0.423772
+  GA & 0.083042
+  GF\% & 0.705010
+  xGF & 0.386643
+  xGA & 0.195763
+  xGF\% & 0.585988
+  SCF & 0.387204
+  SCA & 0.191645
+  SCF\% & 0.585252
+  HDCF & 0.399467
+  HDCA & 0.195659
+  HDCF\% & 0.551392
+  HDGF & 0.442448
+  HDGA & 0.092987
+  HDGF\% & 0.614221
+  On-Ice SH\% & 0.324061
+  On-Ice SV\% & 0.419340
+  PDO & 0.546471
+  Off. Zone Starts & 0.309539
+  Neu. Zone Starts & 0.294922
+  Def. Zone Starts & 0.236777
+  On The Fly Starts & 0.273141
+  Off. Zone Start \% & 0.110750
+  Off. Zone Faceoffs & 0.353819
+  Neu. Zone Faceoffs & 0.314974
+  Def. Zone Faceoffs & 0.262650
+  Off. Zone Faceoff \% & 0.182999
+  \bottomrule
+\end{tabular}
+\end{table}
 
-Cross-Validation Scores:
+Metrics such as GF\%, FF\%, SF\%, PDO, xGF\%, and HDGF\% show a relatively strong correlation with plus/minus (0.705010, 0.553993, 0.568873, 
+0.546471, 0.585988, 0.614221 respectively), which aligns with expectations since many of these metrics also track goal-related contributions. 
+Higher positive correlations indicate metrics that vary similarly to plus/minus, suggesting that they may capture overlapping aspects of
+offensive or defensive performance.
+
+Ridge regression was performed on three different sets of variables: combined metrics, offensive only, and defensive only. The combined
+metrics consisted of Goals For Percentage(GF\%), Scoring Chances Percentage (SF\%), High-Danger Chances while that player is on the ice (HDSC\%),
+combined shooting and save percentage (PDO), Percent of shots that were goals for (On-Ice SH\%), High-Danger Goals for while that player is 
+on the ice (HDGC\%). All of these metrics are similar to plus-minus in that they combine the defensive and offensive factors together.
+The offensive only metrics were CF, FF, SF, GF, SCF, HDCF, HDGF, GF\%, G, A, P/GP, CF\%, SCF\%, and On-Ice SH\%. The defensive only metrics were
+CA, FA, SA, GA, SCF, HDCA, HDGA, On-Ice SV\%. The results can be seen in the table below.
 
@@
 
-Mixed Effects Model:
-
-The significant coefficients for both GF\% and team averages, negative for TeamMeanCF\% and positive for TeamMeanGF\%, imply that 
-team-level factors substantially impact individual plus/minus scores.
-This aligns with the idea that plus/minus is both a team and individual statistic, as team-level scoring factors are significant predictors.
-
 \begin{table}[tbp]
@@
   \midrule
-  0.56777061 & 0.64692431 & 0.73265605 \\ 
-  0.51687324 & 0.62667921 & 0.70862251 \\ 
-  0.49576518 & 0.62263538 & 0.69467537 \\ 
-  0.49576518 & 0.58484799 & 0.65315835 \\ 
-  0.61054685 & 0.69145036 & 0.76239753 \\ 
+  0.5696297 & 0.64692431 & 0.73265605 \\ 
+  0.50125171 & 0.62667921 & 0.70862251 \\ 
+  0.47617049 & 0.62263538 & 0.69467537 \\ 
+  0.48944826 & 0.58484799 & 0.65315835 \\ 
+  0.57586267 & 0.69145036 & 0.76239753 \\ 
   \bottomrule
@@
 \end{table}
+
+Mixed Effects Model:
+
+The significant coefficients for both GF\% and team averages, negative for TeamMeanCF\% and positive for TeamMeanGF\%, imply that 
+team-level factors substantially impact individual plus/minus scores.
+This aligns with the idea that plus/minus is both a team and individual statistic, as team-level scoring factors are significant predictors.
+
+Plus-Minus vs Corsi vs Fenwick
+
+Both Corsi and Fenwick were used as predictors in ridge regression with the variables GF\%, SCF\%, HDCF\%, PDO, On-Ice SH\%, On-Ice SV\%, G, A.
+The results of the cross validation are shown in the table below
+
+\begin{table}[tbp]
+  \caption{Cross Validation Scores}
+  \label{tab:rv}
+\centering
+\begin{tabular}{rrr}
+  \toprule
+Corsi & Fenwick \\ 
+  \midrule
+  0.93898296 & 0.95099819 \\ 
+  0.90054897 & 0.93012992 \\ 
+  0.93915158 & 0.9528202 \\ 
+  0.92841222 & 0.95201592 \\ 
+  0.9325304 & 0.95921829 \\ 
+   \bottomrule
+\end{tabular}
+\end{table}
+
+These values are notibly better that the cross validation scores for plus-minus, regardless of which set of predictors. This indicates
+that both Corsi and Fenwick are significantly better at predicting player effectiveness than plus-minus.
 

### 2024-11-01 code/analysis.py
made some changes to the variables used for rr

@@ -10,9 +10,6 @@ nhl = nhl[nhl['GP_x'] > 25]
 nhl_edge_numeric = nhl_edge.select_dtypes(include=['number'])
 correlation_matrix = nhl_edge_numeric.corr()
 #print(correlation_matrix)
-with open("correlation_matrix.tex", "w") as file:
-    file.write(correlation_matrix.to_latex())
-
 
 import matplotlib.pyplot as plt
 """
@@ -38,7 +35,7 @@ from sklearn.preprocessing import StandardScaler
 # Define target and features (Offensive vs. Defensive)
 # Offensive variables: CF, FF, SF, GF, SCF, HDCF, HDGF, GF%, G, A, P/GP, CF%, SCF%, On-Ice SH%
 # Defensive variables: CA, FA, SA, GA, SCF, HDCA, HDGA, On-Ice SV%
-X = nhl[['GF%', 'SCF%', 'HDCF%', 'PDO', 'On-Ice SH%', 'On-Ice SV%']]
+X = nhl[['GF%', 'SCF%', 'HDCF%', 'PDO', 'CF%', 'HDGF%', 'FF%']]
 y = nhl['+/-']  # Plus/minus as the target variable
 
 # Scale the features for Ridge Regression

### 2024-11-01 manuscript/statspaper.tex
finished all sections

@@
 
-Use this section to describe the data that helps to answer your research
-questions.
-
 The data used to perform this analysis was collected from Natural Stat Trick and the NHL website. The observations in the data
@@
 calliber and were regular NHL players. The columns of the dataset will be combined from the two data sources. The following is a 
-description of each of the columns in the dataset obtained from the Natural Stat Trick website:
+description of each of the columns in the dataset obtained from the Natural Stat Trick website \cite{natural_stat_trick}:
 
@@
 
-Starts
+SH\% - Percentage of Shots for that player's team while that player is on the ice that were Goals. GF*100/SF
 
-Off. Zone Starts - Number of shifts for the player that started with an offensive zone faceoff.
-
-Neu. Zone Starts - Number of shifts for the player that started with an neutral zone faceoff.
-
-Def. Zone Starts - Number of shifts for the player that started with an defensive zone faceoff.
-
-On The Fly Starts - Number of shifts for the player that started during play (without a faceoff).
-
-Off. Zone Start \% - Percentage of starts for the player that were Offensive Zone Starts, excluding Neutral Zone and On The Fly Starts. Off. Zone Starts*100/(Off. Zone Starts+Def. Zone Starts)
-
-Faceoffs
-
-Off. Zone Faceoffs - Number of faceoffs in the offensive zone for which the player was on the ice.
-
-Neu. Zone Faceoffs - Number of faceoffs in the neutral zone for which the player was on the ice.
-
-Def. Zone Faceoffs - Number of faceoffs in the defensive zone for which the player was on the ice.
-
-Off. Zone Faceoff \% - Percentage of faceoffs in the offensive zone for which the player was on the ice, excluding neutral zone faceoffs. Off. Zone Faceoffs*100/(Off. Zone Faceoffs+Def. Zone Faceoffs)
+SV\% - Percentage of Shots against that player's team while that player is on the ice that were not Goals. GA*100/SA
 
@@
 \label{sec:meth}
-
-Use this section to present the methodologies that will generate results by
-analyzing the data.
 
@@
 \subsection{Cross-Validation}
+
 Coupled with ridge regression, cross validation was utilized to assess the predictive power of different groups of variables (offensive, 
@@
 \subsection{Mixed Model Effects}
-In order to separate individual contributions to plus/minus from team-level effects, a mixed-effects model was used. Mixed-effects models 
-allow for random effects, capturing variability at the team level, while fixed effects capture individual-level metrics. Individual metrics 
-like CF\%, GF\%, and others were included that directly measure a player's on-ice performance. The model considered team-level averages 
-(e.g., TeamMeanGF\%), accounting for how a player’s team context might influence their plus/minus.
+
+In order to separate individual contributions to plus/minus from team-level effects, a mixed-effects model was used. 
+The mixed-effects model combines:
+\begin{itemize}
+    \item \textbf{Fixed Effects}, which capture the influence of player-specific variables that directly relate to individual performance, 
+    such as Corsi For Percentage (CF\%), Goals For Percentage (GF\%), and other advanced metrics.
+    \item \textbf{Random Effects}, which account for variability at the team level, recognizing that a player’s plus/minus statistic can be 
+    influenced by the overall performance and style of their team. Including team-level random effects helps control for unobserved team 
+    factors that may affect each player similarly.
+\end{itemize}
+
+To apply this model, individual metrics that describe on-ice performance (e.g., CF\%, GF\%) were used as fixed effects to estimate each player’s 
+contribution to plus/minus. At the same time, team-level averages (e.g., TeamMeanCF\%, TeamMeanGF\%) were included as random effects. This 
+approach distinguishes how much of a player's plus/minus statistic is attributable to their own performance versus the performance of their team.
+By using this mixed-effects approach, we can assess to what extent the plus/minus statistic reflects individual skill as opposed to team 
+strength, thus helping clarify if plus/minus can serve as a reliable individual performance measure.
 
@@
 
-The results of taking the correlation coefficients of plus-minus versus every other column is as follows:
+The results of the correlation analysis and statistical modeling of plus-minus reveal several insights into how this statistic reflects 
+both individual and team-level contributions.
+
+\subsection{Correlation Analysis}
+
+The initial analysis involved calculating the correlation coefficients between plus-minus and various performance metrics. The correlation 
+coefficients are displayed in Table \ref{tab:rv}.
 
 \begin{table}[tbp]
-  \caption{Cross Validation Scores}
+  \caption{Correlation Coefficients with Plus-Minus}
   \label{tab:rv}
 \centering
-\begin{tabular}{rrr}
+\begin{tabular}{rr}
   \toprule
@@
   \midrule
-  GP & 0.232144
-  G & 0.237058
-  A & 0.342538
-  P & 0.314842
-  +/- & 1.000000
-  P/GP & 0.277666
-  EVG & 0.239076
-  EVP & 0.336372
-  PPG & 0.187284
-  PPP & 0.226803
-  SHG & 0.177479
-  SHP & 0.282570
-  OTG & 0.161693
-  GWG & 0.305998
-  S & 0.255841
-  S\% & 0.048282
-  TOI/GP & 0.253234
-  TOI & 0.291711
-  CF & 0.369200
-  CA & 0.210752
-  CF\% & 0.533927
-  FF & 0.373707
-  FA & 0.213121
-  FF\% & 0.553993
-  SF & 0.376144
-  SA & 0.211659
-  SF\% & 0.568873
-  GF & 0.423772
-  GA & 0.083042
-  GF\% & 0.705010
-  xGF & 0.386643
-  xGA & 0.195763
-  xGF\% & 0.585988
-  SCF & 0.387204
-  SCA & 0.191645
-  SCF\% & 0.585252
-  HDCF & 0.399467
-  HDCA & 0.195659
-  HDCF\% & 0.551392
-  HDGF & 0.442448
-  HDGA & 0.092987
-  HDGF\% & 0.614221
-  On-Ice SH\% & 0.324061
-  On-Ice SV\% & 0.419340
-  PDO & 0.546471
-  Off. Zone Starts & 0.309539
-  Neu. Zone Starts & 0.294922
-  Def. Zone Starts & 0.236777
-  On The Fly Starts & 0.273141
-  Off. Zone Start \% & 0.110750
-  Off. Zone Faceoffs & 0.353819
-  Neu. Zone Faceoffs & 0.314974
-  Def. Zone Faceoffs & 0.262650
-  Off. Zone Faceoff \% & 0.182999
+  GP & 0.232 \\ 
+  G & 0.237 \\ 
+  A & 0.343 \\ 
+  P & 0.315 \\ 
+  +/- & 1.000 \\ 
+  P/GP & 0.278 \\ 
+  EVG & 0.239 \\ 
+  EVP & 0.336 \\ 
+  PPG & 0.187 \\ 
+  PPP & 0.227 \\ 
+  SHG & 0.177 \\ 
+  SHP & 0.283 \\ 
+  OTG & 0.162 \\ 
+  GWG & 0.306 \\ 
+  S & 0.256 \\ 
+  S\% & 0.048 \\ 
+  TOI/GP & 0.253 \\ 
+  TOI & 0.292 \\ 
+  CF & 0.369 \\ 
+  CA & 0.211 \\ 
+  CF\% & 0.534 \\ 
+  FF & 0.374 \\ 
+  FA & 0.213 \\ 
+  FF\% & 0.554 \\ 
+  SF & 0.376 \\ 
+  SA & 0.212 \\ 
+  SF\% & 0.569 \\ 
+  GF & 0.424 \\ 
+  GA & 0.083 \\ 
+  GF\% & 0.705 \\ 
+  xGF & 0.387 \\ 
+  xGA & 0.196 \\ 
+  xGF\% & 0.586 \\ 
+  SCF & 0.387 \\ 
+  SCA & 0.192 \\ 
+  SCF\% & 0.585 \\ 
+  HDCF & 0.399 \\ 
+  HDCA & 0.196 \\ 
+  HDCF\% & 0.551 \\ 
+  HDGF & 0.442 \\ 
+  HDGA & 0.093 \\ 
+  HDGF\% & 0.614 \\ 
+  On-Ice SH\% & 0.324 \\ 
+  On-Ice SV\% & 0.419 \\ 
+  PDO & 0.546 \\ 
+  Off. Zone Starts & 0.310 \\ 
+  Neu. Zone Starts & 0.295 \\ 
+  Def. Zone Starts & 0.237 \\ 
+  On The Fly Starts & 0.273 \\ 
+  Off. Zone Start \% & 0.111 \\ 
+  Off. Zone Faceoffs & 0.354 \\ 
+  Neu. Zone Faceoffs & 0.315 \\ 
+  Def. Zone Faceoffs & 0.263 \\ 
+  Off. Zone Faceoff \% & 0.183 \\ 
   \bottomrule
@@
 
-Metrics such as GF\%, FF\%, SF\%, PDO, xGF\%, and HDGF\% show a relatively strong correlation with plus/minus (0.705010, 0.553993, 0.568873, 
-0.546471, 0.585988, 0.614221 respectively), which aligns with expectations since many of these metrics also track goal-related contributions. 
-Higher positive correlations indicate metrics that vary similarly to plus/minus, suggesting that they may capture overlapping aspects of
-offensive or defensive performance.
+Metrics such as GF\%, FF\%, SF\%, PDO, xGF\%, and HDGF\% show moderate to strong positive correlations with plus-minus (0.705, 0.554, 
+0.569, 0.546, 0.586, and 0.614, respectively). This result aligns with expectations, as many of these metrics are related to goal-scoring 
+chances and shot control, which contribute directly to team scoring and subsequently affect plus-minus. Higher correlations among these 
+metrics indicate a similar variation pattern to plus-minus, suggesting that they capture overlapping aspects of a player's offensive and 
+defensive performance.
+
+\subsection{Ridge Regression Analysis of Offensive and Defensive Contributions}
 
 Ridge regression was performed on three different sets of variables: combined metrics, offensive only, and defensive only. The combined 
-metrics consisted of Goals For Percentage(GF\%), Scoring Chances Percentage (SF\%), High-Danger Chances while that player is on the ice (HDSC\%),
-combined shooting and save percentage (PDO), Percent of shots that were goals for (On-Ice SH\%), High-Danger Goals for while that player is 
-on the ice (HDGC\%). All of these metrics are similar to plus-minus in that they combine the defensive and offensive factors together.
-The offensive only metrics were CF, FF, SF, GF, SCF, HDCF, HDGF, GF\%, G, A, P/GP, CF\%, SCF\%, and On-Ice SH\%. The defensive only metrics were
-CA, FA, SA, GA, SCF, HDCA, HDGA, On-Ice SV\%. The results can be seen in the table below.
-
-Offensive/Defensive Combined: Averages around 0.53, indicating that these metrics explain a moderate proportion of the variance in plus/minus.
-Offensive Only: Cross-validation scores are around 0.63, suggesting that offensive metrics alone are somewhat predictive of plus/minus, 
-reflecting the offensive impact.
-Defensive Only: With higher scores around 0.71, defensive metrics seem to have a greater influence on the plus/minus, which aligns with 
-the role of plus/minus in capturing defensive contributions.
-Corsi and Fenwick: Both show very high cross-validation scores, with Fenwick performing slightly better, indicating that these 
-possession-based metrics are strong predictors.
+metrics included GF\%, SF\%, HDSC\%, PDO, On-Ice SH\%, and HDGF\%, metrics that combine both offensive and defensive factors, similar to 
+plus-minus.The offensive only metrics were CF, FF, SF, GF, SCF, HDCF, HDGF, GF\%, G, A, P/GP, CF\%, SCF\%, and On-Ice SH\%. The defensive 
+only metrics were CA, FA, SA, GA, SCF, HDCA, HDGA, On-Ice SV\%. The results can be seen in the table below.
 
@@
 
+The results in Table \ref{tab:rv} reveal:
+- **Combined Metrics**: A moderate cross-validation score around 0.53 suggests that these metrics explain a moderate amount of variance 
+in plus-minus.
+- **Offensive Metrics**: With an average score of 0.63, offensive metrics are predictive of plus-minus to a certain extent, indicating 
+the offensive contributions captured by plus-minus.
+- **Defensive Metrics**: The highest average score (around 0.71) suggests that defensive metrics have a more substantial impact on 
+plus-minus, aligning with the fact that plus-minus also reflects defensive contributions.
+
 Mixed Effects Model:
 
-The significant coefficients for both GF\% and team averages, negative for TeamMeanCF\% and positive for TeamMeanGF\%, imply that 
-team-level factors substantially impact individual plus/minus scores.
-This aligns with the idea that plus/minus is both a team and individual statistic, as team-level scoring factors are significant predictors.
+The results of the mixed effects model can be summarized in the table.
 
-Plus-Minus vs Corsi vs Fenwick
+\begin{table}[tbp]
+  \caption{Mixed Effects Model Results for Plus-Minus Prediction}
+  \label{tab:mixed-effects}
+  \centering
+  \begin{tabular}{lrr}
+    \toprule
+    Predictor & Coefficient & Significance \\ 
+    \midrule
+    GF\% & 1.950 & 0.000 \\
+    CF\% & 0.313 & 0.345 \\ 
+    Team Mean CF\% & -3.111 & 0.000 \\ 
+    Team Mean GF\% & 0.846 & 0.000 \\ 
+    \bottomrule
+\end{tabular}
+\end{table}
+
+The coefficient for individual Corse For percentage is a small, positive effect on plus/minus, though it is not statistically significant 
+(p = 0.345). This suggests that while Corsi percentage (CF\%) has a slight positive association with plus/minus, it may not have a 
+substantial or reliable impact on explaining variance in plus/minus at the individual level. The positive and highly significant 
+coefficient for individual GF\% (p < 0.001) indicates a strong and reliable positive relationship between goals-for percentage and plus/minus. 
+Players with a higher individual GF\% are likely to have a higher plus/minus score, suggesting that goal-scoring and offensive contribution 
+are important for explaining plus/minus. The coefficent for Team Mean CF\% negative, statistically significant coefficient (p < 0.001) 
+indicates that the team’s mean CF\% negatively impacts individual plus/minus. In teams with higher Corsi percentages, individual players 
+might have lower plus/minus scores, possibly due to the distribution of possession-based contributions across the team. The positive 
+coefficient of Team Mean GF\%, significant at p < 0.01, confirms that team-level GF\% positively influences individual plus/minus. This 
+suggests that players benefit from being on teams that are generally good at scoring, supporting the notion that plus/minus partially 
+reflects team-level offensive strength. These results highlight that plus/minus is significantly influenced by both individual offensive 
+performance (e.g., GF\%) and team-level scoring success (e.g., Team Mean GF\%), while possession metrics like CF\% are less predictive of 
+plus/minus, especially at the individual level.
+
+\subsection{Plus-Minus vs Corsi vs Fenwick}
 
@@
 
-What are the main contributions again?
+Like the plus-minus statistic, this project does not come without its limitations. This analysis relies on specific datasets from 
+Natural Stat Trick and includes only a selection of individual and team-based metrics. Other relevant factors like zone entries, exits, or 
+additional situational metrics might provide further insights into player contributions but were not available in this dataset. The model 
+does not capture all nuances of game context, such as player fatigue, line changes, and shifts against specific opponents, which can all 
+influence a player’s plus/minus. The mixed-effects model approximates some of these factors with team-level random effects, but a more 
+complex model might be required for a comprehensive understanding.
 
-What are the limitations of this study?
-
-What are worth pursuing further in the future?
+In this paper, the effectiveness of the plus-minus statistic was evaluated to see if it is an effective predictor of player contribution
+in the National Hockey League. The plus minus does have some value as a predictor of offensive and defensive metrics. However, other
+advanced metrics like Corsi and Fenwick perform much better. The plus-minus is also heavily influenced by a player's team, making it 
+less valuable as an indivual evaluator of player effectiveness. Overall, the NHL should maybe consider phasing out the use of the plus-minus
+statistic and consider more advanced metrics such as Corsi or Fenwick.
 

### 2024-11-02 manuscript/statspaper.tex
fixed error in intro

@@
 
-Use this section to answer three questions:
-Why is the topic important/interesting?
-What has been done on this topic in the literature?
-What is your contribution?
-
 Within the realm of sports, many different statistics are used to determine how good a player is. This can include points,
@@
 calliber and were regular NHL players. The columns of the dataset will be combined from the two data sources. The following is a 
-description of each of the columns in the dataset obtained from the Natural Stat Trick website \cite{natural_stat_trick}:
+description of each of the columns in the dataset obtained from the website \cite{natural_stat_trick}:
 

### 2024-11-20 review/PeerReviewByLucy.tex

\documentclass[a4paper,9pt]{scrartcl}
\usepackage{amssymb, amsmath} % needed for math
\usepackage[utf8]{inputenc}   % this is needed for umlauts
\usepackage[USenglish]{babel} % this is needed for umlauts
\usepackage[T1]{fontenc}      % this is needed for correct output of umlauts in pdf
\usepackage[margin=2.5cm]{geometry} %layout
\usepackage{hyperref}         % hyperlinks
\usepackage{color}
\usepackage{framed}
\usepackage{enumerate}  % for advanced numbering of lists
\usepackage{csquotes}   % for enquote

\newcommand\titletext{Peer-Review of\\"Hockey’s Most Controversial Statistic: An Analysis of the Effectiveness of the Plus-Minus Statistic"}

\title{\titletext}
\author{Lucy Liu}

\hypersetup{
  pdfauthor   = {Martin Thoma},
  pdfkeywords = {peer review},
  pdftitle    = {Lineare Algebra}
}

\usepackage{microtype}

\begin{document}
\maketitle
\section{Introduction}
This is a peer-review of \enquote{Hockey’s Most Controversial Statistic: An Analysis of the Effectiveness of the Plus-Minus Statistic } by Melanie Desroches . The reviewed document
is available under \href{https://github.com/sw24f/final-paper-melpd}{https://github.com/sw24f/final-paper-melpd}.

\section{Summary of the Content}
The paper assesses the validity of the plus-minus statistic in evaluating player contributions in the National Hockey League (NHL). It highlights potential limitations, such as multicollinearity with other metrics and team-level dependencies. Using various statistical approaches, including correlation analysis, ridge regression, cross-validation, and mixed-effects models, the study examines how plus-minus compares to alternative metrics like Corsi and Fenwick. The analysis concludes that plus-minus, while reflective of certain aspects of player performance, may be less effective than Corsi or Fenwick in evaluating individual contributions. This study is insightful, but greater clarity and methodological precision would improve the paper.

\section{Overall Feedback}
While the paper provides an explanation of the plus-minus statistic, there are several places that could use more clarification. These include better organization of results, more detailed explanations of methodological choices, and clearer presentation of results and conclusions. The paper should also address minor inconsistencies in terminology and provide more guidance on the data sources used.

\section{Major Remarks}
The literature review could be expanded to include additional studies that have examined the plus-minus statistic and its alternatives. This could provide a better foundation for the reasoning behind this study and highlight gaps that you hope to fill with your current research.

\subsection{Introduction}
Page 1. The paper evaluates the effectiveness of the plus-minus statistic compared to other metrics like Corsi and Fenwick. While the objectives are clear, the hypotheses are not explicitly stated. Clearly state the hypotheses at the end of the introduction.

\subsection{Methods Section 3.1-3.4}
Pages 6-9. Although each method is described individually, the reason behind selecting these specific methods (Ex. why ridge regression and not another regularization technique) is not fully explained. Add a brief justification for each method and explain why alternatives were not selected.

Additionally, clearly outline the assumptions associated with each statistical method and discuss how these assumptions are checked and addressed within the analysis. Include diagnostic tests and results that assess whether the assumptions of each statistical method are satisfied. For example, provide residual plots or normality tests for mixed-effects models.

\subsection{Discussion}
Page 13.  The discussion section is missing future directions. Discuss what comes next after this study, or some possible new ideas or improvements to explore. Further elaborate on the implications of these findings for the NHL.

\section{Minor Remarks}

\begin{itemize}
    \item Line 196 remove \enquote{we}, paper is written by a single author. 
    \item Terminology Consistency (Methods, Pages 6-9):
    \begin{itemize}
        \item Terms such as \enquote{multicollinearity}, \enquote{collinearity}, “collinearity,” \enquote{correlation} are used interchangeably but may confuse readers. Consistent use of terminology would improve clarity. 
    \end{itemize}
    \item Tables on page 7 and 10 are cut off, and could use captions
    \item Asterisks are showing instead of bold text. Ex. Line 220. **Combined Metrics** 
    \item References are sometimes not fully cited. Ex. Macdonald’s work mentioned without complete reference. 
    \item typos
    \begin{itemize}
        \item Line 9: \enquote{valaue} should be \enquote{value}
        \item Line 13: \enquote{refered} should be \enquote{referred}
        \item Line 19: \enquote{it’s} should be \enquote{its}
        \item Line 20: \enquote{independece} should be \enquote{independence}
        \item Line 29: \enquote{colinearity} should be \enquote{collinearity}
        \item Line 37: \enquote{truely} should be \enquote{truly}
        \item Line 38: \enquote{hereforth} should be \enquote{henceforth}
        \item Line 39: \enquote{preforming} should be \enquote{performing}
        \item Line 50: \enquote{ensure} should be \enquote{to ensure}
        \item Line 50: \enquote{calliber} should be \enquote{caliber}
        \item Line 141: \enquote{uncovential} should be \enquote{unconventional}
        \item Line 151: \enquote{makes since} should be \enquote{makes sense}
        \item Line 177: \enquote{R-sqared} should be \enquote{R-squared}
        \item Line 229: \enquote{Corse} should be \enquote{Corsi}
        \item Line 236: \enquote{coefficient for Team Mean CF\% negative} should be \enquote{coefficient for Team Mean CF\% is negative}
        \item Line 252: \enquote{notibly} should be \enquote{notably}
        \item Line 269: \enquote{indivual} should be \enquote{individual} 
    \end{itemize}
\end{itemize}

\end{document}

### 2024-11-20 manuscript/statspaper.tex

@@
 \begin{abstract}
-Here is the abstract.  
+The goal of this paper is to evaluate the plus minus statistic as a metric for determining the effectiveness of a player.
+This involves analysing the relationship between a players plus-minus rating and other offensive and defensive ratings,
+the effect of a players team performance on their plus-minus, and comparing plus-minus to other similar statistics. Based
+on the analysis performed, the plus minus does have some value as a predictor of offensive and defensive metrics. However, other
+advanced metrics like Corsi and Fenwick perform much better. The plus-minus is also heavily influenced by a player's team, making it 
+less valuable as an indivual evaluator of player effectiveness.
 \end{abstract}
@@
 advanced metrics like Corsi and Fenwick perform much better. The plus-minus is also heavily influenced by a player's team, making it 
-less valuable as an indivual evaluator of player effectiveness. Overall, the NHL should maybe consider phasing out the use of the plus-minus
+less valuable as an indivual evaluator of player effectiveness. Overall, the NHL should consider phasing out the use of the plus-minus
 statistic and consider more advanced metrics such as Corsi or Fenwick.

### 2024-11-21 manuscript/statspaper.tex
fixed spelling mistakes

@@
 important topic, as fans, coaches, players, and general managers often want to get a better understanding of how an individual is
-impacting their team. It is one thing to watch the players to determine how good a player is, often refered to as the "eye test" but 
+impacting their team. It is one thing to watch the players to determine how good a player is, often referred to as the "eye test" but 
 with so many players, it is hard to quantify this. Enter the plus-minus statistic. Plus-minus is generally calculated by adding all the points
@@
 
-While the plus-minus statistic is a great idea in theory, it does not come without it's weaknesses. The biggest drawback to this
-statistic is its issues with independece. Take ice hockey for example. Many ice hockey players tend to play on "lines", meaning the same three 
+While the plus-minus statistic is a great idea in theory, it does not come without its weaknesses. The biggest drawback to this
+statistic is its issues with independence. Take ice hockey for example. Many ice hockey players tend to play on "lines", meaning the same three 
 forwards tend to play together and the same two defensemen usually play together. As a result, the performance of one player is highly dependent 
@@
  \cite[Daniel spent 92\% of his playing time with Henrik, the highest percentage of any other player combination where both players have
-played over 700 minutes. Because of this high colinearity between the twins, it is difficult to separate the individual effect that each 
+played over 700 minutes. Because of this high collinearity between the twins, it is difficult to separate the individual effect that each 
 player has on the net goals scored on the ice.]{Macdonald_2012} Many critics of the plus-minus argue that plus-minus is more of a team 
@@
 
-The goal of this project will be to shed light on this controversial statistic and determine if plus-minus is truely an effective predictor
-to individual contribution, specifically in the National Hockey League (hereforth refered to as the NHL). Is the plus-minus a good evaluator 
-of offensive and defensive output? Is the plus-minus more reflective of how well a player is doing than how well a player is preforming? 
+The goal of this project will be to shed light on this controversial statistic and determine if plus-minus is truly an effective predictor
+to individual contribution, specifically in the National Hockey League (henceforth refered to as the NHL). Is the plus-minus a good evaluator 
+of offensive and defensive output? Is the plus-minus more reflective of how well a player is doing than how well a player is performing? 
 Are there better alternatives to evaluating individual player contribution? These are the questions that will be answered from this analysis.
@@
 is from all players that played in the National Hockey League from the 2021-2022, 2022-2023, and 2023-2024 seasons. Only players
-that have played in over 25 games were used in the analysis. This is ensure that the players being evaluated were all of similar 
+that have played in over 25 games were used in the analysis. This is to ensure that the players being evaluated were all of similar 
 calliber and were regular NHL players. The columns of the dataset will be combined from the two data sources. The following is a 
@@
 The goal of this dataset was to provide a broader scope of variables that evaluate a players performance. By introducing more
-advanced and uncovential statistics, there are more factors to consider in evaluating a player's effectiveness on the ice.
+advanced and unconventional statistics, there are more factors to consider in evaluating a player's effectiveness on the ice.
 
@@
 There is a high correlation between different shot related statistics and point related statistics. For example, there is a 
-0.826371 correlation between Goals and Shots. This makes since because in order to score a goal, the player needs to shoot first.
+0.826371 correlation between Goals and Shots. This makes sense because in order to score a goal, the player needs to shoot first.
 Correlation analysis can be used to identify the strength and direction of the relationship between various advanced metrics and 
@@
 computing a model's predictive accuracy each time. In this case, the data was split into five subsets. In each iteration, one subset was
-kept as the test set and the other four were used to train the ridge regression model. The R-sqared value was reported for each iteration
+kept as the test set and the other four were used to train the ridge regression model. The R-squared value was reported for each iteration
 to measure how well offensive, defensive, and possession-based metrics (like Corsi and Fenwick) predict plus/minus. This process ensures 
@@
 The results in Table \ref{tab:rv} reveal:
+
 - **Combined Metrics**: A moderate cross-validation score around 0.53 suggests that these metrics explain a moderate amount of variance 
 in plus-minus.
+
 - **Offensive Metrics**: With an average score of 0.63, offensive metrics are predictive of plus-minus to a certain extent, indicating 
 the offensive contributions captured by plus-minus.
+
 - **Defensive Metrics**: The highest average score (around 0.71) suggests that defensive metrics have a more substantial impact on 
@@
 
-The coefficient for individual Corse For percentage is a small, positive effect on plus/minus, though it is not statistically significant 
+The coefficient for individual Corsi For percentage is a small, positive effect on plus/minus, though it is not statistically significant 
 (p = 0.345). This suggests that while Corsi percentage (CF\%) has a slight positive association with plus/minus, it may not have a 
@@
 Players with a higher individual GF\% are likely to have a higher plus/minus score, suggesting that goal-scoring and offensive contribution 
-are important for explaining plus/minus. The coefficent for Team Mean CF\% negative, statistically significant coefficient (p < 0.001) 
+are important for explaining plus/minus. The coefficent for Team Mean CF\% is negative, statistically significant coefficient (p < 0.001) 
 indicates that the team’s mean CF\% negatively impacts individual plus/minus. In teams with higher Corsi percentages, individual players 
@@
 
-These values are notibly better that the cross validation scores for plus-minus, regardless of which set of predictors. This indicates
+These values are notably better that the cross validation scores for plus-minus, regardless of which set of predictors. This indicates
 that both Corsi and Fenwick are significantly better at predicting player effectiveness than plus-minus.
@@
 advanced metrics like Corsi and Fenwick perform much better. The plus-minus is also heavily influenced by a player's team, making it 
-less valuable as an indivual evaluator of player effectiveness. Overall, the NHL should consider phasing out the use of the plus-minus
+less valuable as an individual evaluator of player effectiveness. Overall, the NHL should consider phasing out the use of the plus-minus
 statistic and consider more advanced metrics such as Corsi or Fenwick.

### 2024-12-04 code/analysis.py
added assumptions and made small changes to the first model

@@ -1,3 +1,4 @@
+'''
 import pandas as pd
 
 nhl_edge = pd.read_csv('data\cleaned_nhl_edge.csv')
@@ -25,8 +26,150 @@ plus_minus_corr = nhl.select_dtypes(include=['number']).corr()['+/-']
 print(plus_minus_corr)
 
 print(nhl.columns)
-# how plus/minus can be employed to assess offensive and defensive contribution with ridge regression/regular regression
+'''
+import pandas as pd
+import matplotlib.pyplot as plt
+from statsmodels.stats.outliers_influence import variance_inflation_factor
+from statsmodels.stats.diagnostic import het_breuschpagan
+from scipy.stats import shapiro
+import seaborn as sns
+import statsmodels.api as sm
+from sklearn.linear_model import Ridge, RidgeCV
+from sklearn.model_selection import train_test_split, cross_val_score
+import numpy as np
+from sklearn.preprocessing import StandardScaler
 
+# Load datasets
+nhl_edge = pd.read_csv('data/cleaned_nhl_edge.csv')
+nst = pd.read_csv('data/cleaned_nst.csv')
+nst_indiv = pd.read_csv('data/cleaned_nst_indiv.csv')
+nst = nst.drop_duplicates(subset='Player', keep='first')
+nst_indiv = nst_indiv.drop_duplicates(subset='Player', keep='first')
+
+# Merge datasets
+nhl = pd.merge(nhl_edge, nst, on='Player', how='inner')  # Merge nhl_edge with nst
+nhl = pd.merge(nhl, nst_indiv, on='Player', how='inner')  # Merge with nst_indiv
+
+nhl['Team'] = nhl['Team_x'].combine_first(nhl['Team_y'])
+nhl['Position'] = nhl['Position_x'].combine_first(nhl['Position_y'])
+nhl['GP'] = nhl['GP_x'].combine_first(nhl['GP_y'])
+nhl['TOI'] = nhl['TOI_x'].combine_first(nhl['TOI_y'])
+
+# Clean up unnecessary columns
+nhl.drop(columns=['Team_x', 'Team_y', 'Position_x', 'Position_y', 'GP_x', 'GP_y', 'TOI_x', 'TOI_y'], inplace=True)
+
+# Filter for players with more than 25 games played
+nhl = nhl[nhl['GP'] > 25]
+
+print(nhl.columns)
+# Select predictors and target
[…]

### 2024-12-04 code/data_cleaning.py
added the individual stats

@@ -172,3 +172,13 @@ csv_file_path = 'data/cleaned_nhl_edge.csv'
 nst = clean_columns(nst, ['SF%', 'GF%', 'SCF%', 'HDCF%','HDGF%', 'On-Ice SH%', 'On-Ice SV%', 'PDO', 'Off. Zone Start %', 'Off. Zone Faceoff %'])
 print(nst.dtypes)
 nst.to_csv('data/cleaned_nst.csv', index=False)
+
+nst_indiv = pd.read_csv('data\Player Season Totals_individual - Natural Stat Trick.csv')
+
+print(nst_indiv.dtypes)
+
+nst_indiv['IPP'].fillna(0, inplace= True)
+nst_indiv['SH%'].fillna(0, inplace=True)
+nst_indiv['Faceoffs %'].fillna(0, inplace=True)
+
+nst_indiv.to_csv('data/cleaned_nst_indiv.csv', index=False)

### 2024-12-04 manuscript/statspaper.tex
added to discussion and intro

@@
 Within the realm of sports, many different statistics are used to determine how good a player is. This can include points,
-assists, time played, etc. But is there just one valaue that can be used to determine the effectiveness of a player? This is an 
+assists, time played, etc. But is there just one value that can be used to determine the effectiveness of a player? This is an 
 important topic, as fans, coaches, players, and general managers often want to get a better understanding of how an individual is
-impacting their team. It is one thing to watch the players to determine how good a player is, often referred to as the "eye test" but 
+impacting their team. It is one thing to watch the players to determine how good a player is, often referred to as the "eye test", but 
 with so many players, it is hard to quantify this. Enter the plus-minus statistic. Plus-minus is generally calculated by adding all the points
 scored by their team while they were playing and subtracting by points scored by the opposition while they were playing. The idea is that 
-if a player has a generally positive impact on their team, they will have a highly positive plus-minus. The two main sports where this
-statistic is used is basketball and ice hockey.
+if a player has a generally positive impact on their team, they will have a highly positive plus-minus. In this project, plus-minus will
+be evaluated in the context of ice hockey in the National Hockey League (henceforth refered to as the NHL).
+
+The use of the plus-minus originated in ice hockey by the NHL team Montréal Canadiens \cite{plus_minus_rating}. The league
+started to keep track of plus-minus in 1956. The NHL defines plus-minus as "a team's goal differential while a particular player 
+is on the ice, excluding power play goals for and against but including empty net situations. All the skaters on the ice receive a plus 
+or minus when an even-strength goal or shorthanded goal is scored depending on which team scored" \cite{nhl_website}. This means
+that when a goal is scored, as long as there are the same number of players from each team on the ice, the team that scored gets plus one
+and the other team gets a minus one.
 
 While the plus-minus statistic is a great idea in theory, it does not come without its weaknesses. The biggest drawback to this
-statistic is its issues with independence. Take ice hockey for example. Many ice hockey players tend to play on "lines", meaning the same three 
+statistic is its issues with independence. Many ice hockey players tend to play on "lines", meaning the same three 
 forwards tend to play together and the same two defensemen usually play together. As a result, the performance of one player is highly dependent 
@@
 
-Medium Danger Scoring Chances - a scoring chance with a score of exactly 2.
-
-MDCF - Count of Medium Danger Scoring Chances for that player's team while that player is on the ice.
-
-MDCA - Count of Medium Danger Scoring Chances against that player's team while that player is on the ice.
-
-MDCF\% - Percentage of total Medium Danger Scoring Chances while that player is on the ice that are for that player's team. MDCF*100/(MDCF+MDCA)
-
-Medium Danger Goals - goals generated from Medium Danger Scoring Chances
-
-MDGF - Count of Goals off of Medium Danger Scoring Chances for that player's team while that player is on the ice.
-
-MDGA - Count of Goals off of Medium Danger Scoring Chances against that player's team while that player is on the ice.
-
-MDGF\% - Percentage of Medium Danger Goals while that player is on the ice that are for that player's team. MDGF*100/(MDGF+MDGA)
-
-Low Danger Scoring Chances - a scoring chance with a score of 1 or less. Does not include any attempts from the attacking team's neutral or defensive zone.
-
-LDCF - Count of Low Danger Scoring Chances for that player's team while that player is on the ice.
-
-LDCA - Count of Low Danger Scoring Chances against that player's team while that player is on the ice.
-
-LDCF\% - Percentage of total Low Danger Scoring Chances while that player is on the ice that are for that player's team. LDCF*100/(LDCF+LDCA)
-
-Low Danger Goals - goals generated from Low Danger Scoring Chances
-
-LDGF - Count of Goals off of Low Danger Scoring Chances for that player's team while that player is on the ice.
-
-LDGA - Count of Goals off of Low Danger Scoring Chances against that player's team while that player is on the ice.
-
-LDGF\% - Percentage of Low Danger Goals while that player is on the ice that are for that player's team. LDGF*100/(LDGF+LDGA)
-
-PDO
-
 SH\% - Percentage of Shots for that player's team while that player is on the ice that were Goals. GF*100/SF
@@
 SV\% - Percentage of Shots against that player's team while that player is on the ice that were not Goals. GA*100/SA
+
+Goals - Goals scored by the player, outside of the shootout.
+
+Assists - Any assist by the player.
+
+First Assists - Primary assists by the player.
+
+Second Assists - Secondary assists by the player.
+
+Total Points - Goals scored and assists by the player, outside of the shootout.
+
+IPP - Individual Point Percentage, the percentage of goals for that player's team while that player is on the ice that the 
+player earned a point on.
+
+Shots - Any shot attempt on net (goals and shots on net) by the player, outside of the shootout.
+
+SH\% - Percentage of Shots by the player that were Goals. Goals*100/Shots
+
+iCF - Any shot attempt (goals, shots on net, misses and blocks) by the player, outside of the shootout.
+
+iFF - Any unblocked shot attempt (goals, shots on net and misses) by the player, outside of the shootout.
+
+iSCF - Any scoring chance by the player, outside of the shootout.
+
+iHDCF - Any high danger scoring chance by the player, outside of the shootout.
+
+Rush Attempts - Any rush shot attempt (goals, shots on net, misses and blocks) by the player, outside of the shootout.
+
+Rebounds Created - Any shot attempt (shots on net, misses and blocks) that results in a rebound shot attempt.
+
+Penalties Drawn - Number of penalties committed against the player.
+
+Giveaways - Number of unforced turnovers made by the player.
+
+Takeaways - Number of times the player takes the puck away from the opposition.
+
+Hits - Number of hits made by the player.
+
+Hits Taken - Number of hits taken by the player.
+
+Shots Blocked - Number of opposition shot attempts blocked by the player.
 
@@
 
+In order for fans, players, and sports management to have a better understanding of what make a player "effective", you have to 
+summarize and quantify a large number of variables into one condensed number. That is, understandably, a tall order. While the plus-minus
+lays the ground-work for getting a better understanding of player effectiveness, there is more at play. General managers of sports teams
+often want to make the best team possible, getting players that will have a positive impact. Often times, people put a lot of emphasis on
+direct offensive output, such as points, to determine a players contribution. But there is so much more to consider, such as defensive skills,
+generating chances to score, etc, while still accounting for team effects. While plus-minus may not be the solution, the next step in the 
+world of sports statistics is to find a way to quantify the concept of "player effectiveness".
+
 \bibliography{refs}

### 2024-12-05 manuscript/statspaper.tex
changed the format of the tables to fit

@@
 \begin{abstract}
+
 The goal of this paper is to evaluate the plus minus statistic as a metric for determining the effectiveness of a player.
@@
 less valuable as an indivual evaluator of player effectiveness.
+
 \end{abstract}
@@
 MacDonald provides a perfect example of this with the Henrik and Daniel Sedin, Swedish twins who played together for the Vancouver Canucks. 
- \cite[Daniel spent 92\% of his playing time with Henrik, the highest percentage of any other player combination where both players have
+"Daniel spent 92\% of his playing time with Henrik, the highest percentage of any other player combination where both players have
 played over 700 minutes. Because of this high collinearity between the twins, it is difficult to separate the individual effect that each 
-player has on the net goals scored on the ice.]{Macdonald_2012} Many critics of the plus-minus argue that plus-minus is more of a team 
+player has on the net goals scored on the ice."\cite{Macdonald_2012} Many critics of the plus-minus argue that plus-minus is more of a team 
 statistic, since it is heavily influenced by team dynamics rather than individual contributions. Furthermore, numerous confounding variables, 
 such as the quality of the opponent and situational factors, make the calculation less reliable. As a result of these criticisms, alternatives 
-to the plus-minus statistic have been developed, such as Corsi or Fenwick, which is calculated using shot attempts rather than goals.
+to the plus-minus statistic have been developed, such as Corsi or Fenwick, which is calculated using shot attempts rather than goals. Brian 
+MacDonald introduces an adjusted plus-minus statistic that is based on weighted least squares regression\cite{Macdonald_2012}.
 
@@
 The main issue with the data is multicollinearity, where independent variables are highly correlated. A correlation matrix was created 
-with some of the more basic nhl statistics to highlight this issue, shown in Table \ref{tab:correlation_matrix}.
+with some of the more basic nhl statistics to highlight this issue, shown in Table \ref{tab:correlation_matrix2}.
 
@@
   \centering
+  \scriptsize
   \caption{Correlation Matrix of Variables}
-  \input{correlation_matrix.tex}
-  \label{tab:correlation_matrix}
+  \input{correlation_matrix2.tex}
+  \label{tab:correlation_matrix2}
 \end{table}
@@
 \centering
+\small
 \begin{tabular}{rr}
@@
   PDO & 0.546 \\ 
-  Off. Zone Starts & 0.310 \\ 
-  Neu. Zone Starts & 0.295 \\ 
-  Def. Zone Starts & 0.237 \\ 
-  On The Fly Starts & 0.273 \\ 
-  Off. Zone Start \% & 0.111 \\ 
-  Off. Zone Faceoffs & 0.354 \\ 
-  Neu. Zone Faceoffs & 0.315 \\ 
-  Def. Zone Faceoffs & 0.263 \\ 
-  Off. Zone Faceoff \% & 0.183 \\ 
   \bottomrule

### 2024-12-06 manuscript/statspaper.tex
added some of the changes tot he methods and results

@@
 applies to hockey metrics like Corsi, Fenwick, scoring chances, etc. since these statistics often interact or overlap in measuring aspects 
-of performance.
+of performance. Ridge regression is appropriate because the data satisfies the model assumptions of independence, homoscedasticity, 
+and linearity. It is ideal for this project because it effectively manages multicollinearity, retains all variables for interpretability, 
+and improves predictive accuracy. Other regularization techniques like Lasso or Elastic Net are better suited for sparse models or when 
+variable selection is a priority, which is not the project's focus.
 
@@
 By using this mixed-effects approach, we can assess to what extent the plus/minus statistic reflects individual skill as opposed to team 
-strength, thus helping clarify if plus/minus can serve as a reliable individual performance measure.
+strength, thus helping clarify if plus/minus can serve as a reliable individual performance measure. The model was tested to ensure that the
+assumptions of linearity, independence, normality, homoscedasticity, and random sampling were not violated.
 
@@
 The initial analysis involved calculating the correlation coefficients between plus-minus and various performance metrics. The correlation 
-coefficients are displayed in Table \ref{tab:rv}.
+coefficients are displayed in Table \ref{tab:rv1}.
 
@@
   \caption{Correlation Coefficients with Plus-Minus}
-  \label{tab:rv}
+  \label{tab:rv1}
 \centering
@@
   S\% & 0.048 \\ 
-  TOI/GP & 0.253 \\ 
   TOI & 0.292 \\ 
@@
 Ridge regression was performed on three different sets of variables: combined metrics, offensive only, and defensive only. The combined 
-metrics included GF\%, SF\%, HDSC\%, PDO, On-Ice SH\%, and HDGF\%, metrics that combine both offensive and defensive factors, similar to 
-plus-minus.The offensive only metrics were CF, FF, SF, GF, SCF, HDCF, HDGF, GF\%, G, A, P/GP, CF\%, SCF\%, and On-Ice SH\%. The defensive 
-only metrics were CA, FA, SA, GA, SCF, HDCA, HDGA, On-Ice SV\%. The results can be seen in the table below.
+metrics included GF\%, SF\%, PDO, Takeaways, and SCF\%, metrics that combine both offensive and defensive factors, similar to 
+plus-minus.The offensive only metrics were CF, FF, SF, GF, SCF, HDCF, HDGF, GF\%, G, A, P/GP, CF\%, SCF\%, Rush Attempts, and On-Ice SH\%. 
+The defensive only metrics were SA, GA, SCF, HDCA, HDGA, On-Ice SV\%, Hits, Shots Blocked, and Penalties Drawn. The results can be seen in the table below.
 
@@
   \caption{Cross Validation Scores}
-  \label{tab:rv}
+  \label{tab:rv2}
 \centering
@@
 
-The results in Table \ref{tab:rv} reveal:
+The results in Table \ref{tab:rv2} reveal:
 
@@
     \midrule
-    GF\% & 1.950 & 0.000 \\
-    CF\% & 0.313 & 0.345 \\ 
+    GF\% & 1.947 & 0.000 \\
+    CF\% & 0.446 & 0.478 \\
+    SCF\% & -0.109 & 0.848 \\
     Team Mean CF\% & -3.111 & 0.000 \\ 
     Team Mean GF\% & 0.846 & 0.000 \\
+    Team SCF\% & 3.127 & 0.001 \\ 
     \bottomrule
@@
 The coefficient for individual Corsi For percentage is a small, positive effect on plus/minus, though it is not statistically significant 
-(p = 0.345). This suggests that while Corsi percentage (CF\%) has a slight positive association with plus/minus, it may not have a 
-substantial or reliable impact on explaining variance in plus/minus at the individual level. The positive and highly significant 
-coefficient for individual GF\% (p < 0.001) indicates a strong and reliable positive relationship between goals-for percentage and plus/minus. 
+(p = 0.478). This suggests that while Corsi percentage (CF\%) has a slight positive association with plus/minus, it may not have a 
+substantial or reliable impact on explaining variance in plus/minus at the individual level. Similarly, Scoring Chances For percentage 
+(SCF\%) also does not have a significant relationship with plus/minus (p = 0.848).The positive and highly significant coefficient
+for individual GF\% (p < 0.000) indicates a strong and reliable positive relationship between goals-for percentage and plus/minus. 
 Players with a higher individual GF\% are likely to have a higher plus/minus score, suggesting that goal-scoring and offensive contribution 
-are important for explaining plus/minus. The coefficent for Team Mean CF\% is negative, statistically significant coefficient (p < 0.001) 
-indicates that the team’s mean CF\% negatively impacts individual plus/minus. In teams with higher Corsi percentages, individual players 
-might have lower plus/minus scores, possibly due to the distribution of possession-based contributions across the team. The positive 
-coefficient of Team Mean GF\%, significant at p < 0.01, confirms that team-level GF\% positively influences individual plus/minus. This 
-suggests that players benefit from being on teams that are generally good at scoring, supporting the notion that plus/minus partially 
-reflects team-level offensive strength. These results highlight that plus/minus is significantly influenced by both individual offensive 
-performance (e.g., GF\%) and team-level scoring success (e.g., Team Mean GF\%), while possession metrics like CF\% are less predictive of 
-plus/minus, especially at the individual level.
+are important for explaining plus/minus.
+
+The coefficent for Team Mean CF\% is negative, statistically significant coefficient (p < 0.000) indicates that the team’s mean CF\% 
+negatively impacts individual plus/minus. In teams with higher Corsi percentages, individual players might have lower plus/minus scores, 
+possibly due to the distribution of possession-based contributions across the team. The positive coefficient of Team Mean GF\%, 
+significant at p < 0.01, confirms that team-level GF\% positively influences individual plus/minus. This suggests that players benefit 
+from being on teams that are generally good at scoring, supporting the notion that plus/minus partially reflects team-level offensive 
+strength. Finally, the positive and significant coefficient for Team SCF (p=0.001) indicates that teams generating more scoring chances 
+positively affect individual plus/minus. This highlights the role of team offensive strength in influencing this statistic.
+
+These results confirm that plus/minus is heavily influenced by team-level performance metrics (e.g., Team Mean GF\% and Team SCF) 
+and less so by individual possession metrics like CF\% and SCF\%. The significant relationship between GF\% and plus/minus supports 
+the notion that offensive contributions and team success are key drivers of this statistic, reinforcing that plus/minus is a 
+team-oriented measure rather than solely reflective of individual performance.
 

### 2024-12-06 code/analysis.py
fixed ridge

@@ -201,7 +201,7 @@ y_pred = ridge.predict(X_test)
 '''
 
 # Offensive alone
-X_o = nhl[['CF', 'FF', 'SF', 'GF', 'SCF', 'HDCF', 'HDGF', 'GF%', 'G', 'A', 'P/GP', 'CF%', 'SCF%', 'On-Ice SH%', 'Rush Attempts']]
+X_o = nhl[['SF', 'GF', 'SCF', 'HDCF', 'HDGF', 'GF%', 'G', 'A', 'P/GP', 'On-Ice SH%', 'Rush Attempts']]
 y_o = nhl['+/-']  # Plus/minus as the target variable
 
 # Scale the features for Ridge Regression
@@ -211,8 +211,15 @@ X_o_scaled = scaler.fit_transform(X_o)
 # Split the data into training and test sets
 X_train_o, X_test_o, y_train_o, y_test_o = train_test_split(X_o_scaled, y_o, test_size=0.2, random_state=42)
 
+alphas = np.logspace(-6, 6, 200)  # Range of alphas to test
+ridge_cv = RidgeCV(alphas=alphas, scoring='neg_mean_squared_error', cv=5)  # 5-fold CV
+ridge_cv.fit(X_train, y_train)
+
+best_alpha = ridge_cv.alpha_
+print(f"\nOptimal Lambda (alpha): {best_alpha}")
+
 # Fit Ridge Regression Model
-ridge_o = Ridge(alpha=1.0)  # alpha is the regularization strength (lambda in ridge regression)
+ridge_o = Ridge(alpha=best_alpha)  # alpha is the regularization strength (lambda in ridge regression)
 ridge_o.fit(X_train_o, y_train_o)
 
 # Evaluate the model using cross-validation
@@ -223,7 +230,7 @@ print("Cross-Validation Scores (Offensive):", cv_scores_o)
 y_pred_o = ridge_o.predict(X_test_o)
 
 # Defensive output
-X_d = nhl[['SA', 'GA', 'SCF', 'HDCA', 'HDGA', 'On-Ice SV%', 'Hits', 'Shots Blocked']]
+X_d = nhl[['SA', 'GA', 'SCF', 'HDCA', 'HDGA', 'On-Ice SV%', 'Hits', 'Shots Blocked', 'Penalties Drawn']]
 y_d = nhl['+/-']  # Plus/minus as the target variable
 
 # Scale the features for Ridge Regression
@@ -234,7 +241,13 @@ X_scaled_d = scaler.fit_transform(X_d)
[…]

### 2024-12-06 response/response.tex
did all but one response to comment

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{listings}
\usepackage{rotating, graphicx}
\usepackage{booktabs, natbib}
\usepackage[hyphens]{url}
% \usepackage [english]{babel}
\usepackage{amsmath, amsbsy, amsthm, epsfig, epsf, psfrag, graphicx,
  amssymb, enumerate, bm}
\usepackage{enumitem}

\usepackage{color}

\newcommand{\jy}[1]{\textcolor{red}{JY #1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}

\usepackage[colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=blue]{hyperref}

\newenvironment{comment}%
{\begin{quotation}\noindent\small\it\ignorespaces%
  }{\end{quotation}}

\begin{document}

\begin{center}
  {\Large\bf Response Letter to Reviewers' Comments}
\end{center}

\section{Summary}

I would like to that Lucy Liu for her thoughtful, constructive
comments. The manuscript has been revised accordingly with the
following major changes:

\begin{enumerate}
\item
  Following the reviewer's comments on expanding the literature
  review, a new paragraph was added to the introduction that
  expanded on other works that have been done. This involved reviewing
  alternatives to the plus-minus and works that other people
  within the sports statistics world have done to address this.
  This also included addressing some downsides to these works and
  why this project is necessary.

\item
  In response to the reviewer's comment on clearly stating the 
  hypothesis, the hypothesis was added to the last paragraph of the 
  introduction that asked the over-arching question that is being
  addressed in the paper. This set up the work that is being done
  in the paper and what the readers can hope to take from this work.

\item
  The reasons that each of the models used were expanded. The need
  for the assumptions to varified was added to the methods section.
  The reviewer did make the suggestion of adding the plots to varify 
  that the assumptions were satisfied. While I did make plots to make
  sure all assumptions were varified, they were not included in the final
  paper due to space concerns. Also, I did not want to take attention
  away from the discussion at hand. There were additions made to the 
  ridge regression section explaining why ridge was chosen over other 
  regularization techniques.

\item 
  An extra paragraph was added to the end of the discussion to explain
  future directions. The reason for this study was re-iterated as well as
  explaining how it can be used in the NHL. Additional comments were added
  to explain what can be done in the future based on this study and the
  work of other people within the field.

\end{enumerate}

Point-by-point responses are as follows, with the reviewer/editor's comments in {\it italic}.

\section{Response to Each Comment}

\begin{comment}
  The literature review could be expanded to include additional studies that 
  have examined the plus-minus statistic and its alternatives. This could provide 
  a better foundation for the reasoning behind this study and highlight gaps
  that you hope to fill with your current research.
\end{comment}

This was a good note to improve the paper. While I initially mentioned Brian MacDonald's 
work and impliedthat other people had proposed alternatives to the plus-minus, I did not 
include enough detail. Including more literature helps the reader to have a better understanding
of what work is already being done. It also allows me to explain why my study is 
important and what it contributed to this conversation. A paragraph better explaining 
the literature review was added as a fourth paragraph in the introduction.

\begin{comment}
  The paper evaluates the effectiveness of the plus-minus statistic compared 
  to other metrics like Corsi and Fenwick. While the objectives are clear, the 
  hypotheses are not explicitly stated. Clearly state the hypotheses at the end 
  of the introduction.
\end{comment}

In the first draft, the overall goal of the project was stated as well as the 
questions that would be answered in the paper. At the request of the reviewer,
the hypothesis of the project was stated in the second to last sentence of the
introduction. Hopefully this helps the readers to understand the over-arching 
question that is being answered in this project.

\begin{comment}
  Although each method is described individually, the reason behind selecting these 
  specific methods (Ex.why ridge regression and not another regularization technique) 
  is not fully explained. Add a brief justification for each method and explain why 
  alternatives were not selected.

  Additionally, clearly outline the assumptions associated with each statistical 
  method and discuss how these assumptions are checked and addressed within the 
  analysis. Include diagnostic tests and results that assess whether the assumptions 
  of each statistical method are satisfied. For example, provide residual plots 
  or normality tests for mixed-effects models.
\end{comment}

\begin{comment}
   The discussion section is missing future directions. Discuss what comes next after 
   this study, or some possible new ideas or improvements to explore. Further elaborate 
   on the implications of these findings for the NHL.
\end{comment}

\bibliographystyle{chicago}
\bibliography{growth}

%\atColsBreak{\pagediscards}
\end{document}

### 2024-12-06 code/data_cleaning.py
added notes about the other stats

@@ -63,6 +63,34 @@ Off. Zone Faceoffs - Number of faceoffs in the offensive zone for which the play
 Neu. Zone Faceoffs - Number of faceoffs in the neutral zone for which the player was on the ice.
 Def. Zone Faceoffs - Number of faceoffs in the defensive zone for which the player was on the ice.
 Off. Zone Faceoff % - Percentage of faceoffs in the offensive zone for which the player was on the ice, excluding neutral zone faceoffs. Off. Zone Faceoffs*100/(Off. Zone Faceoffs+Def. Zone Faceoffs)
+Goals - Goals scored by the player, outside of the shootout.
+Assists - Any assist by the player.
+First Assists - Primary assists by the player.
+Second Assists - Secondary assists by the player.
+Total Points - Goals scored and assists by the player, outside of the shootout.
+IPP - Individual Point Percentage, the percentage of goals for that player's team while that player is on the ice that the player earned a point on. Total Points/Goals For
+Shooting
+Shots - Any shot attempt on net (goals and shots on net) by the player, outside of the shootout.
+SH% - Percentage of Shots by the player that were Goals. Goals*100/Shots
+iCF - Any shot attempt (goals, shots on net, misses and blocks) by the player, outside of the shootout.
+iFF - Any unblocked shot attempt (goals, shots on net and misses) by the player, outside of the shootout.
+iSCF - Any scoring chance by the player, outside of the shootout.
+iHDCF - Any high danger scoring chance by the player, outside of the shootout.
+Rush Attempts - Any rush shot attempt (goals, shots on net, misses and blocks) by the player, outside of the shootout.
+Rebounds Created - Any shot attempt (shots on net, misses and blocks) that results in a rebound shot attempt.
+Penalties
+PIM - Penalty minutes assessed to the player.
+Total Penalties - Total number of penalties taken by the player
+Minor - Number of minor penalties taken by the player.
+Major - Number of major penalties taken by the player.
+Misconduct - Number of 10 minute or game misconducts taken by the player.
+Penalties Drawn - Number of penalties committed against the player.
+Real-time
+Giveaways - Number of unforced turnovers made by the player.
+Takeaways - Number of times the player takes the puck away from the opposition.
+Hits - Number of hits made by the player.
+Hits Taken - Number of hits taken by the player.
+Shots Blocked - Number of opposition shot attempts blocked by the player.
 """
 
 import pandas as pd

### 2024-12-06 response/response.tex
finished hopefully

@@
 
+While in the first draft, the reason that each method was being used was explained,
+the reviewer is correct that there is a need for additional explanation of why each
+model works in the given context. A few sentences were added to the begining of the 
+Ridge Regression section in Methods to explain how the models were used. Additionally,
+the assumptions of the models needed to be addressed. While I did test that the 
+assumptions were not violated, images were not included. Due to the length of the paper,
+I did not want to include images to keep the paper from being incredibly long. Also, I felt
+that it would not benefit the discussion to include a longer sectiona about verifying the
+models' assumptions. A few sentences were added to the end of each of the models' sections 
+to verify that the assumptions were met.
+
 \begin{comment}
@@
 
+The reviewer is correct, the paper needs a better ending that encapsulates the work 
+done. The initial last papragraph explained what was accomplished in this study and
+breifly mentioned the implications of this work but failed to remind the readers why
+it is important and what is can be done moving forward. An extra paragraph was added 
+to the end of the Discussion section to explain why this study was important, what can 
+be done with the information in this paper, and what next steps can be taken.
+
 \bibliographystyle{chicago}

### 2024-12-06 manuscript/statspaper.tex
added based on comments

@@
 statistic, since it is heavily influenced by team dynamics rather than individual contributions. Furthermore, numerous confounding variables, 
-such as the quality of the opponent and situational factors, make the calculation less reliable. As a result of these criticisms, alternatives 
-to the plus-minus statistic have been developed, such as Corsi or Fenwick, which is calculated using shot attempts rather than goals. Brian 
-MacDonald introduces an adjusted plus-minus statistic that is based on weighted least squares regression\cite{Macdonald_2012}.
+such as the quality of the opponent and situational factors, make the calculation less reliable.
+
+As a result of these criticisms, alternatives to the plus-minus statistic have been developed. Many of these alternatives are either different 
+metrics or adjusted plus-minus\cite{plus_minus_rating}. Two popular examples are Corsi or Fenwick, which are calculated using shot attempts 
+rather than goals. Another popular metric is expected goals, which is based on the probability of a goal being scored depending on the shot.
+Brian MacDonald introduced an adjusted plus-minus statistic that is based on weighted least squares regression\cite{Macdonald_2012}. However,
+many of these alternatives are more complicated to calculate, which can be confusing to the average fan. Due to the fact that plus-minus
+is easy to interpret and calculate, it has remained in use despite its controversial background.
 
@@
 of offensive and defensive output? Is the plus-minus more reflective of how well a player is doing than how well a player is performing? 
-Are there better alternatives to evaluating individual player contribution? These are the questions that will be answered from this analysis.
+Are there better alternatives to evaluating individual player contribution? Is the plus-minus an effective predictor to individual contribution? 
+These are the questions that will be answered from this analysis.
 
@@
 
+Ridge regression is used in this project to evaluate the relationship between various performance metrics and plus-minus, while addressing 
+multicollinearity among predictor variables. This regularization technique is particularly valuable in this context because many hockey 
+performance metrics, such as Scoring Chances For percentage (SCF\%) and Goals For percentage (GF\%), are often highly correlated.
 Due to the presence of correlation among variables, ridge regression was performed in order to identify how plus/minus can be employed 
@@
 direct offensive output, such as points, to determine a players contribution. But there is so much more to consider, such as defensive skills,
-generating chances to score, etc, while still accounting for team effects. While plus-minus may not be the solution, the next step in the 
-world of sports statistics is to find a way to quantify the concept of "player effectiveness".
+generating chances to score, etc, while still accounting for team effects. Many have tried to do this, but lack the interpretability and 
+simplicity of the plus-minus. While plus-minus may not be the solution, the next step in the world of sports statistics is to find a way to 
+quantify the concept of "player effectiveness" in a way that includes the simplicity of the plus-minus without its controversial nature.
 

### 2024-12-06 code/analysis.py
cleaned up the code

@@ -1,32 +1,4 @@
-'''
-import pandas as pd
 
-nhl_edge = pd.read_csv('data\cleaned_nhl_edge.csv')
-nst = pd.read_csv('data\cleaned_nst.csv')
-nhl = pd.merge(nhl_edge, nst, on='Player', how='inner')  # 'inner' merge keeps only players present in both datasets
-nhl = nhl[nhl['GP_x'] > 25]
-# Identify collinearity
-#print(nhl_edge.describe())
-# Select only numerical columns
-nhl_edge_numeric = nhl_edge.select_dtypes(include=['number'])
-correlation_matrix = nhl_edge_numeric.corr()
-#print(correlation_matrix)
-
-import matplotlib.pyplot as plt
-"""
-# Scatter plot of shots vs goals
-plt.scatter(nhl_edge['+/-'], nhl_edge['G'])
-plt.xlabel('Plus/Minus')
-plt.ylabel('Goals')
-plt.title('Plus/Minus vs Goals')
-plt.show()
-"""
-
-plus_minus_corr = nhl.select_dtypes(include=['number']).corr()['+/-']
-print(plus_minus_corr)
-
-print(nhl.columns)
-'''
 import pandas as pd
 import matplotlib.pyplot as plt
 from statsmodels.stats.outliers_influence import variance_inflation_factor
@@ -62,60 +34,10 @@ nhl.drop(columns=['Team_x', 'Team_y', 'Position_x', 'Position_y', 'GP_x', 'GP_y'
 nhl = nhl[nhl['GP'] > 25]
 
 print(nhl.columns)
-# Select predictors and target
+
 X = nhl[['GF%', 'PDO', 'SF%', 'Takeaways', 'SCF%']]
-y = nhl['+/-']  # Target variable
-'''
-from sklearn.preprocessing import PolynomialFeatures, StandardScaler
-from sklearn.linear_model import Ridge
-from sklearn.pipeline import Pipeline
-from sklearn.model_selection import train_test_split
-from sklearn.metrics import mean_squared_error, r2_score
+y = nhl['+/-'] 
 
-# Split your data into train and test sets
-X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.2, random_state=42)
-
-# Create a pipeline to combine polynomial feature generation, scaling, and ridge regression
-degree = 2  # Adjust the degree based on your observations
[…]
